\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}
\item 


%% -- [Total 11]
CT4 A2013–4
8
During a football match, the referee can caution players if they commit an offence by
showing them a yellow card. If a player commits a second offence which the referee
deems worthy of a caution, they are shown a red card, and are sent off the pitch and
take no further part in the match. If the referee considers a particularly serious
offence has been committed, he can show a red card to a player who has not
previously been cautioned, and send the player off immediately.
The football team manager can also decide to substitute one player for another at any
point in the match so that the substituted player takes no further part in the match.
Due to the risk of a player being sent off, the manager is more likely to substitute a
player who has been shown a yellow card. Experience shows that players who have
been shown a yellow card play more carefully to try to avoid a second offence.
The rate at which uncautioned players are shown a yellow card is 1/10 per hour.
The rate at which those players who have already been shown a yellow card are
shown a red card is 1/15 per hour.
The rate at which uncautioned players are shown a red card is 1/40 per hour.
The rate at which players are substituted is 1/10 per hour if they have not been shown
a yellow card, and 1/5 if they have been shown a yellow card.
(i) Sketch a transition graph showing the possible transitions between states for a
given player.

(ii) Write down the compact form of the Kolmogorov forward equations,
specifying the generator matrix.

A football match lasts 1.5 hours.
(iii) Solve the Kolmogorov equation for the probability that a player who starts the
match remains in the game for the whole match without being shown a yellow
card or a red card.

(iv) Calculate the probability that a player who starts the match is sent off during
the match without previously having been cautioned.

Consider a match that continued indefinitely rather than ending after 1.5 hours.
(v)
CT4 A2013–5
(a) Derive the probability that in this instance a player is sent off without
previously having been cautioned.
(b) Explain your result.

%% -- [Total 12]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

8
(i)
1/40
Uncautioned
1/10
1/15
1/5
1/10
Substituted
Page 10
Booked
/Yellow
Sent Off
/Red %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(ii)
d
P ( t ) = P ( t ) A
dt
where generator matrix
⎛ − 9 / 40 1/10 1/ 40 1/10 ⎞
⎜
⎟
0
− 4 / 15 1/15 1/ 5 ⎟
A = ⎜
⎜ 0
0
0
0 ⎟
⎜
⎟
0
0
0 ⎠
⎝ 0
In order of states {U, Y, R, S}
(iii)
d
9
P UU ( t ) = − P UU ( t )
dt
40
⎛ 9 ⎞
⎛ 9 ⎞
P UU ( t ) = exp ⎜ − t ⎟ + const = exp ⎜ − t ⎟ as looking for probability on
⎝ 40 ⎠
⎝ 40 ⎠
pitch throughout match.
At end t = 3/2 so require exp(−27/80) = 71.36%.
(iv)
Prob[sent off without being booked] =
= ∫
3/2
s = 0
exp( −
3/2
1
∫ s = 0 P UU ( s ). 40 ds
9
1
s ). ds
40 40
3/2
1 ⎛
⎡ 1
⎛ 27 ⎞ ⎞
⎛ 9 ⎞ ⎤
= ⎢ − exp ⎜ − s ⎟ ⎥ = ⎜ 1 − exp ⎜ − ⎟ ⎟ =0.03183
9 ⎝
⎝ 80 ⎠ ⎠
⎝ 40 ⎠ ⎦ 0
⎣ 9
(v)
(a)
EITHER
The upper limit of the integral tends to infinity
so result becomes 1/9.
OR
We need
Pr[sent off directly]/Pr[leaves state U]
=
(b)
1/ 40 1
=
9 / 40 9
This is the ratio of the transition rate to “straight to sent off”
Page 11 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
to the total transition rate out of state U.
This was one of the more demanding questions on the paper, and a high proportion of
candidates struggled to get past part (ii). In part (i) the transition rates were not required. In
part (iii) the question asked candidates to solve an equation, so for full credit the equation
had to be written down, and the method of solution described. In part (v) candidates who
used the rationale in (b) to do the calculation in (a) scored full credit.
Candidates who interpreted the question in a manner not intended, but instead combined the
categories “sent off” and “substituted” were not penalised. This interpretation leads to a
three state solution for parts (i) and (ii) as follows.
(i)
Uncautioned
1/10
Booked
/Yellow
1/15
1/5 +
1/15
1/10
+
1/40
Not playing
d
P ( t ) = P ( t ) A
dt
(ii)
where generator matrix
⎛ − 9
⎜ 40
⎜
A = ⎜ 0
⎜
⎜
⎜ ⎜ 0
⎝
1
1 ⎞
10 8 ⎟
⎟
− 4 4 ⎟
15 15 ⎟
⎟
0
0 ⎟
⎟
⎠
In order of states {U, Y, N}

This was given full credit in parts (i), (ii) and (iii). The answer to part (iii) is the same as for
the four-state solution. Credit was given in parts (iv) and (v) for following this alternative
through correctly.
Page 12 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
