\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

%% © Institute and Faculty of Actuaries

1 Describe the differences between deterministic and stochastic models.



%%%%%%%%%%%%%%%%%%%%%%%%%
%% 1
\begin{itemize}
\item A stochastic model is one that recognises the random nature of the input components.
A model that does not contain any random component is deterministic in nature.
\item In a deterministic model, the output is determined once the set of fixed inputs and the
relationships between them have been defined.
By contrast, in a stochastic model the output is random in nature. The output is only a
snapshot or an estimate of the characteristics of the model for a given set of inputs.
A deterministic model is really just a special (simplified) case of a stochastic model.
\item A deterministic model will give one set of results of the relevant calculations for a single
scenario; a stochastic model will be run many times with the same input and gives
distributions of the relevant results for a distribution of scenarios
\item The results for a deterministic model can often be obtained by direct calculation.
The results of stochastic models often require Monte Carlo simulation, although some
stochastic models can have an analytical solution.
\item Correlations can be important in stochastic models as they indicate when the behaviour of
one variable is associated with that of another.
\item Stochastic models are more complex and more difficult to interpret than deterministic models
and so require more expertise, expense and computer power.
\end{itemize}

%%Not all the points listed above were required for full marks. Credit was also given for
%%sensible points not included in the above list.


\end{document}
