\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

9
10
A life office compared the mortality of its policyholders in the age range 30 to 60 years inclusive with a set of mortality rates prepared by the Continuous Mortality Investigation (CMI). The mortality of the life office policyholders was higher than
the CMI rates at ages 30–35, 38–41, 45–50 and 54–59 years inclusive, and lower than the CMI rates at all other ages in the age range.
\begin{enumerate}[(i)]
\item (i) Perform two tests of the null hypothesis that the underlying mortality of the life office policyholders is represented by the CMI rates.
\item (ii) Comment on your results from part (i).
\item (ii) Explain the problem which duplicate policies cause in the context of the CMI mortality investigations.
\end{itemize}


%% -- [Total 13]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

9
(i)
Signs Test
Under the null hypothesis that the underlying mortality of the life office policyholders
is the same as the CMI mortality,
the number of positive deviations is distributed Binomial( m ,0.5)
THEN EITHER ALTERNATIVE 1 (NORMAL APPROXIMATION)
Here we have m = 31, so as m > 20 we can use the Normal approximation, that the
number of positive deviations is distributed Normal ( m /2, m /4).
⎛ 31 31 ⎞
the number of positive deviations is Normal ⎜ , ⎟ .
⎝ 2 4 ⎠
In this case we have 22 positive deviations.
The z -score corresponding to 22 is
22 − 15.5 6.5
=
= 2.33
2.78
7.75
OR Using a continuity correction
21.5 − 15.5
6
=
= 2.16
2.78
7.75
\begin{itemize}
\item Using a 2-tailed test, we reject the null hypothesis at the 5\% level of
significance if | z |>1.96.
\item The z -score corresponding to 22 is
\item Since 2.33 (or 2,16) > 1.96 we reject the null hypothesis.
OR ALTERNATIVE 2 (EXACT CALCULATION)
\item In this case we have 22 positive deviations.
\item The probability of observing exactly 22 positive deviations is 0.009388
OR
\item The probability of observing ≥ 22 positive deviations is 0.014725
\item Using a 2-tailed test, we reject the null hypothesis at the 5\% level of significance if the probability is <0.025
\item Since 0.014725 < 0.025 we reject the null hypothesis.
\item Grouping of Signs Test
Define the test statistic:
G = Number of groups of positive deviations.
%%--- Page 13 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{itemize}
THEN EITHER ALTERNATIVE 1 (NORMAL APPROXIMATION)
Since m = 31 (which is ≥ 20), we can use a Normal approximation as follows:
⎛ n 1 ( n 2 + 1) ( n 1 n 2 ) 2 ⎞
,
G ~ Normal ⎜
3 ⎟
⎜ n + n
⎟ .
⎝ 1 2 ( n 1 + n 2 ) ⎠
In this case m = 31, n 1 = 22 and n 2 = 9.
Thus G ~ Normal ( 7.10,1.32 ) .
We have 4 groups of positive signs.
The z -score corresponding to 4 is
4 − 7.10 − 3.10
=
= − 2.70
1.15
1.32
Using a 1-tailed test, we reject the null hypothesis at the 5% level of significance if z
< −1.645.
Since −2.70 < −1.645 we reject the null hypothesis.
OR ALTERNATIVE 2 USING TABLE IN GOLD BOOK
\begin{itemize}
\item m = total number of deviations = 31
\item $n_1$ = number of positive deviations = 22
\item $n_2$ = number of negative deviations = 9
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We want k * the largest k such that
x
⎛ n − 1 ⎞⎛ n + 1 ⎞ ⎛ n + n ⎞
k = ∑ ⎜ 1 ⎟⎜ 2 ⎟ / ⎜ 1 2 ⎟ < 0.05
t ⎠ ⎝ n 1 ⎠
t = 1 ⎝ t − 1 ⎠⎝
We have 4 groups of positive signs.
The test fails at the 5\% level if G ≤ k *
From the table in the Gold Book k * = 4
Since G is not greater than this, we reject the null hypothesis

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%---- Page 14 ----- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(ii)
The life office’s rates are, overall, different from the CMI rates (actually they are higher).
Additional tests are needed to examine the magnitude of the difference between the two sets of rates.

The shape of the life office’s mortality rates is also rather different from the CMI schedule, and this might require further investigation,
OR
The Grouping of Signs test suggests clumping of the deviations.
It is possible that the difference between the shape of the two sets of rates is so small
in magnitude as to be negligible.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(iii)
\begin{itemize}
\item We can no longer be sure that we are observing a collection of independent claims.
\item It is quite possible that two distinct death claims are the result of the death of the same
life.
\item The effect of this is to increase the variance of the number of claims,
by a factor which may depend on age.
\item This may affect tests based on standardised deviations.
\end{itemize}
%%Answers to this question were very disappointing, especially part (i) . The two tests were often performed in a rather cursory way, with important steps being missed out. In part (i) the null hypothesis only needed stating once. Common errors were to work with only 30 ages or, more seriously, to use eight ages (i.e. to treat each run of consecutive ages of the same sign as a single age).

\end{document}
