\documentclass[a4paper,12pt]{article}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}



%\usepackage{bigints}

\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}
\setcounter{MaxMatrixCols}{10}
\begin{document}
\begin{enumerate}

%%- Question 5 
\item Claims X each year from a portfolio of insurance policies are normally distributed
with mean $\theta$ and variance $\tau^2$. Prior information is that $\theta$ is normally distributed with
known mean \mu and known variance \sigma2.
Aggregate claims over the last n years have been xi for i = 1 to n, and you should
assume that these are independent.
\begin{enumerate}[(a)]
\item (i) Derive the posterior distribution of $\theta$. 
\item (ii) Write down the Bayesian estimate of $\theta$ under quadratic loss. 
\item (iii) Show that the estimate in your answer to part (ii) can be expressed in the
form of a credibility estimate, including statement of the credibility factor Z.
\end{enumerate}

\item % Question 6

\begin{enumerate}
\item (i) Explain what is meant by a saturated model. 
\item(ii) State the definition of the scaled deviance in a fitting under generalised
linear modelling. 
\item (iii) (a) Define both Pearson and deviance residuals.
(b) Explain how these two types of residuals are generally different.
(c) State in which case they are the same. 
\end{enumerate}
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage


Q5 (i) Given
f(\theta) 
2
2 2
exp ( ) exp 1
2 2
\theta  \mu
  
\sigma \sigma
(\theta2  2\mu\theta)
and
p(x\theta) 
2
2
2 2
1 1
exp ( ) exp 1 ( 2 )
2 2
n n
i
i
i i
x x
 
 \theta
   \theta  \theta
\tau \tau  
 2
2
exp 1 ( 2 )
2
 n\theta  nx\theta
\tau
1
=
  n
i
i
x nx

 
 
 
 ,
we want
p(\thetax)  p(x\theta) p(\theta)
 2
2 2 2 2
exp 1
2 2
   n  \theta   \mu  nx  \theta         \sigma \tau   \sigma \tau  
 exp  \tau2  n\sigma2
2\sigma2\tau2 \theta2  2 \mu\tau2  nx\sigma2
\tau2  n\sigma2

 

 
\theta
 

 








2 2 2 2 2
2 2 2 2 exp
2
n nx
n
 \tau  \sigma  \mu\tau  \sigma    \theta   
 \sigma \tau  \tau  \sigma  
 \thetax ~
  2 2 2 2
2 2 2 2 2 2 N n x,
n n n
 \tau \sigma \sigma \tau 
 \mu  
 \tau  \sigma \tau  \sigma \tau  \sigma 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% – September 2015 – Examiners’ Report
Page 6
(ii) The posterior mean (the point estimator under quadratic loss) is
E(\thetax) =
  2 2
2 2 2 2
n x
n n
\tau \sigma
\mu 
\tau  \sigma \tau  \sigma
(iii) E(\thetax) = (1  Z) \mu + Zx
where
Z =
  2
2 2
n
n
\sigma
\tau  \sigma
= 2
2
n
n \tau

\sigma
is the credibility factor. Hence E(\thetax) can be expressed in the form of a
credibility estimate.
Again well prepared candidates who had learnt the relevant bookwork scored
very well on this question. Some candidates attempted to “fudge” the result or
only quoted the result, making no attempt to derive it, and hence scored
poorly.
\end{document}
