\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

\begin{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%

3
Define what it means for a random variable to belong to an exponential
family.
[1]
Show that if a random variable has the exponential distribution it belongs to an
exponential family.
[3]
[Total 4]
A statistician is told that one of two dice has been chosen and rolled, and he is told the
result of the roll. One dice is a conventional dice, but the other has three sides
numbered 2 and three sides numbered 4. If the statistician correctly identifies the dice
he wins a prize of 1.
(i) Determine the total number of decision functions available to the statistician.
[2]
(ii) (a)
Identify the most natural candidate for the decision function.
(b)
Calculate the expected payoff for this function assuming that each of
the two dice are equally likely to be chosen.
[3]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
1
(i)
A random variable Y belongs to an exponential family if the pdf of Y can be
written in the form
⎡ y θ − b ( θ )
⎤
f ( y ; θ , φ ) = exp ⎢
− c ( y , φ ) ⎥
⎣ a ( φ )
⎦
Where a, b and c are functions.
(ii)
Suppose that the parameter of the exponential distribution Y is λ. Then
f ( y ) = λ exp ( −λ y )
= exp [ log λ − λ y ]
⎡ λ y − log λ ⎤
= exp ⎢
− 1 ⎥ ⎦
⎣
Which is of the required form with
θ=λ
a ( φ ) = − 1
b ( θ ) = log θ
c ( y , φ ) = 0
Alternative solution: θ = −λ; a(φ) = 1; b(θ) = −log(−θ); c(y,φ) = 0
This question was answered well.
Page 3
%%%%%%%%%%%%%%%%%%%%%%%%%
 (Statistical Methods) – April 2012 – Examiners’ Report
2
(i)
The decision function must nominate a choice of die for each potential
outcome from the observation.
There are 6 possible outcomes from the die roll and hence 2 × 2 × 2 × 2 × 2 ×
2 = 64 possible decision functions.
(ii)
The most natural candidate is to nominate the conventional die on rolls of
1,3,5,6 and the special die on rolls of 2 or 4.
The expected payoff from this approach is:
2 ⎞
⎛ 4
0.5 × ⎜ × 1 + × 0 ⎟ + 0.5 × 1 = 0.83333
6 ⎠
⎝ 6
Candidates who understood what a decision function is scored well. However, many
candidates struggled to make any headway with this question. For part (i) some candidates
observed that if the dice roll is 1,3,5 or 6 it is obvious that you must chose the conventional
dice. Therefore a choice is only needed on a roll of a 2 or a 4 giving a total of 2 × 2 = 4
functions. This was given full credit if carefully explained.
