\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

\item (i) Explain the census approximation for calculating the exposed to risk between
any two census dates. 
A mortality investigation bureau has collected the following information on number
of policies in-force each year from different companies.
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
Age	&	Year	&	Company A	&	Company B	&	Company C	\\ \hline
54	&	2011	&	3,400	&	1,250	&	5,780	\\ \hline
54	&	2012	&	3,350	&	1,450	&	5,500	\\ \hline
54	&	2013	&	3,000	&	1,500	&	6,010	\\ \hline
55	&	2011	&	3,250	&	1,190	&	6,000	\\ \hline
55	&	2012	&	3,390	&	1,300	&	5,960	\\ \hline
55	&	2013	&	3,100	&	1,440	&	6,030	\\ \hline
56	&	2011	&	3,270	&	1,150	&	5,950	\\ \hline
56	&	2012	&	3,020	&	1,300	&	5,980	\\ \hline
56	&	2013	&	2,950	&	1,500	&	5,990	\\ \hline
\end{tabular}
\end{center}

\begin{itemize}
\item Company A has provided in-force policy data as at the beginning of each calendar
year using age nearest birthday.
%% CT4 S2014–3 PLEASE TURN OVER
\item Company B has provided in-force policy data as at the financial year closing date
(which was 31 March in each year) using age last birthday.
\item Company C has provided in-force policy data as at the end of each calendar year
using age next birthday.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item (ii) Calculate the contribution to central exposed to risk for lives aged 55 last
birthday for the calendar year 2012 for each of the companies. 
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\newpage
3 (i) In survival investigations, population counts will only be available at census
dates.
Define Px,t to be the number of lives under observation, aged x last birthday, at
any time t and suppose that we have the values of Px,t only if t is a census date.
We require the exposed to risk, c
Ex , over the interval between the first census
and the last.
This is
2
1
,
t
c
x x s
t
E   P ds , where t1 and t2 are the two census dates.
To evaluate this, we usually assume that Px,s is linear between census dates.
%%Subject CT4 (Models Core Technical) – September 2014 – Examiners’ Report
%%Page 5
If the censuses are one year apart this leads to the trapezium approximation:
  , 1 , 2
1( )
2
c
Ex  Px t  Px t . [(ii) 
\begin{itemize}
    \item Company A
Assume birthdays evenly distributed across calendar years.
Age 55 last = 0.5 * age 55 nearest + 0.5 * age 56 nearest.
The required exposed to risk is
1 (3,390 3,100 3,020 2,950) 3,115.
4
   
\item Company B
Data are based on age last birthday so no age adjustment needed.
Assuming population varies linearly between census dates, then
population on 1 January 2012 = 31,300 11,190 1,272.5
4 4
 
population on 1 January 2013 = 31,440 11,300 1,405
4 4
 
The required exposed to risk is then
3 1 (1,272.5 1,300) 9 1 1,300 1,405 1,335.9375
12 2 12 2
   
\item Company C
Age 55 last birthday is equivalent to age 56 next birthday.
Assume that data for 31 December in year t apply to 1 January in
year t + 1.
The required exposed to risk is then
1 (5,950 5,980) 5,965.
2
  [6]
[Total 8]
In part (i) equations were not required, explanations and diagrams are acceptable instead.
\end{itemize}
%% Candidates’ answers to this question varied considerably. Partial %% credit was given in part (ii) for a range of alternative 
%% approximations. A common error in part (ii) was to use the wrong %% age when averaging (in particular supposing that age 55 last 
%% could be obtained by averaging age 54 nearest and age 55 nearest).

%% - Subject CT4 (Models Core Technical) 
%% – September 2014 – Examiners’ Report
%% - Page 6
\end{document}
