\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

\begin{enumerate}


4 A manufacturing company is analysing its accident record. The accidents fall into two categories:
   Minor – dealt with by first aider. Average cost £50.
 Major – hospital visit required. Average cost £1,000.
The company has 1,000 employees, of which 180 are office staff and the rest work in the factory.

The analysis shows that 10\% of employees have an accident each year and 20\% of  accidents are major. It is assumed that an employee has no more than one accident in
a year.
\begin{enuerate}[(i)]
\item (i) Determine the expected total cost of accidents in a year. [2]
On further analysis it is discovered that a member of office staff has half the probability of having an accident relative to those in the factory.
\item (ii) Show that the probability that a given member of office staff has an accident in a year is 0.0549. [3]
\item (iii) Determine the probability that a randomly chosen employee who has had an accident is office staff. [2]
\end{enumerate}


Q4
(i)
Let A = Accident
(ii)
Expected cost of an accident= Average Cost(Major)*P(Major|A)
+ Average Cost(Minor)*P(Minor|A)
= 0.2*1000 + 0.8*50 =£240 [1]
Expected total cost of accidents = £240 *1000 * 0.1 = £24,000 [1]
Let O = Office Staff &F = Factory Staff. Then P(A|O) = 0.5 P(A|F)
(iii)
P(A) = P(A|O)P(O)+P(A|F)P(F) = P(A|O)*0.18 + 2*P(A|O)*0.82 [2]
P(A|O) = 0.1 / (0.18+.82*2) = 0.0549 [1]
P(O|A) =
P ( A | O ) P ( O ) 0.0549*0.18
=
= 0.099.
P ( A )
0.1
[2]
[TOTAL 7]
Part (i) was well answered with most candidates giving a fully or partially
correct answer. Answers in parts (ii) and (iii) were generally poor; some
candidates were confused with the conditional probabilities and considered
conditioning on the wrong event.

\end{document}
