\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

\begin{enumerate}
[Total 3]
An insurance company has a portfolio of 10,000 policies. Based on past data the
company estimates that the probability of a claim on any one policy in a year is 0.003.
It assumes no policy will generate more than one claim in a year.
(i)
(ii)
Determine the approximate probability of more than 40 claims from the
portfolio of 10,000 policies in a year. [4]
Determine an approximate equal-tailed interval into which the number of
claims per year will fall with probability 0.95. 
In practice 42 claims were received in a particular year. A Director of the company
complains about the range of estimates in part (ii) being wrong.
(iii)
CT3 S2016–2
Comment on the Director’s complaint.

[Total 8]4
Consider two portfolios, A and B, of insurance policies and denote by X A the number
of claims received in portfolio A and by X B the number of claims received in portfolio
B during a calendar year. The observed numbers of claims received during the last
calendar year are 134 for portfolio A and 91 for portfolio B. X A and X B are assumed
to be independent and to have Poisson distributions with unknown parameters  A and
 B .
Determine an approximate 99% confidence interval for the difference  A   B . You
may use an appropriate normal distribution.
[4]
%%%%%%%%%%%%%%%%%%%%%%%%%
Q3
(i)
Number of claims, N ~ Bin(10000,0.003)
By CLT number of claims approximately N(30, 29.91)

Continuity correction applies
 N − 30 40.5 − 30 
P ( N > 40 ) = P ( N > 40.5 ) = P 
>

29.91 
 29.91
= 1 − Φ ( 1.920 ) = 1 − 0.973 = 0.027
(ii)
95% interval is 30 ± Z 0.975 σ = 30 ± 1.96*5.469 = ( 19.28, 40.72 )


Page 3Subject CT3 (Probability and Mathematical Statistics Core Technical) – September 2016 – Examiners’ Report
(iii)
The probability that the result in any year will lie in the interval is 0.95 so
there is a 5% probability that the company will see a result outside that range.

[Total 8]
Performance here was generally good, but there were some common errors.
In part (i) reference to the CLT was often omitted, as was the continuity
correction (or it was applied in the wrong direction). In part (iii) many
candidates did not identify the main point, about the 95% coverage of the
interval.
Q4
Since the sample size for each portfolio is one, we have that β ˆ A = 134 and β ˆ B = 91 .

Using the normal approximation we find:
β ˆ A − β ˆ B = X A − X B ~ N ( β A − β B , β A + β B )

The confidence interval is then given by
134 − 91 ± 2.5758 134 + 91 = 43 ± 2.5758 × 15 = [ 4.4, 81.6 ]

[Total 4]
Performance in this question was mixed. Many candidates seemed unsure
as to what is the correct size of each sample (portfolio).

