\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

A3
A motor insurer s No Claims Discount system uses the following levels of discount
{0%, 25%, 40%, 50%}. Following a claim free year a policyholder moves up one
discount level (or remains on 50% discount). If the policyholder makes one (or more)
claims in a year they move down one level (or remain at 0% discount).
The insurer estimates that the probability of making at least one claim in a year is 0.1
if the policyholder made no claims the previous year, and 0.25 if they made a claim
the previous year.
New policyholders should be ignored.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(i) Explain why the system with state space {0%, 25%, 40%, 50%} does not form
a Markov chain.
[2]
(ii) (a)
Show how a Markov chain can be constructed by the introduction of
additional states.
(b)
Write down the transition matrix for this expanded system, or draw its
transition diagram.
[4]
(iii) Comment on the appropriateness of the current No Claims Discount system.
[2]
[Total 8]
A3
(i)
AB
t p 0
AA AB
t p 0 . t . dt
lim
dt
0
AB
t p 0 (1
AB
t dt p 0
AB
t p 0
dt
BA
t . dt
BC
t . dt
AA AB
t p 0 . t
BD
t . dt )
AB BA
t p 0 ( t
o ( dt )
BC
t
BD
t )
This is not a Markov chain because it does not possess the Markov property,
that is transition probabilities do not depend only on the current state.
Specifically, if you are in the 25% discount level, the transition probability to
state 0% is 0.25 if a claim was made last year and 0.1 if the previous year was
claim free.
(ii)
(a)
Split the 25% and 40% discount states to include whether the previous
year was claim free.
New state space:
0% discount
25%NC (no claim last year)
25%C (at least one claim last year)
40%NC (no claim last year)
40%C (at least one claim last year)
50%
Page 4Subject CT4
Models Core Technical
April 2006
Examiners Report
(b)
0.9
0.75
0%
25% 40%
NC NC
0.1
0.25
50%
0.75
0.75
0.1
0.25
0.9
25% 40%
C C
0.9
0.1
0.25
New state
(iii)
0%
25%C
25%NC
40%C
40%NC
50%
0%
0.25
0.25
0.1
0
0
0
25%C
0
0
0
0.25
0.1
0
25%NC
0.75
0
0
0
0
0
40%C
0
0
0
0
0
0.1
40%NC
0
0.75
0.9
0
0
0
50%
0
0
0
0.75
0.9
0.9
In theory, the insurer should just use 2 NCD states according to whether the
policyholder made a claim in the previous year. This is because the company
believes the claims frequency is the same for drivers who have not made a
claim for 1, 2, 3 years (i.e. it remains at 0.1 whether the driver has been
claims-free for 1 or 10 years).
However there may be other reasons for adopting this scale:
Marketing or competitive pressures.
It may discourage the policyholder from making small claims, or
encourage careful driving, to preserve their discount.
General comments:
The following, more general comments about the appropriateness of an NCD
model also received credit:
It is appropriate to award a no-claims discount because there is empirical
evidence that drivers who have made a recent claim are more likely to
make a further claim.
More factors should be taken into account (with a suitable example such
as how long the policyholder has been driving).
Page 5Subject CT4
