\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chng%%-- Page}
%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}
\setcounter{MaxMatrixCols}{10}



\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CT6 A2006 24
(i)
Derive the autocovariance and autocorrelation functions of the AR(1) process
X t = X t
where
(ii)
1
+ e t
< 1 and the e t form a white noise process.


The time series Z t is believed to follow a ARIMA(1, d, 0) process for some value of d. The time series $Z_t ( k )$ is obtained by differencing k times and the sample autocorrelations, $\{r_i : i = 1, , 10\}$, are shown in the table below for
various values of k.
\begin{center}
\begin{tabular}{cc|cc|cc}
	&	k = 0	&	k=1	&	k = 2	&	k = 3	&	k = 4	\\ \hline
r 1	&	100\%	&	100\%	&	83\%	&	3\%	&	45\%	\\ \hline
r 2	&	100\%	&	100\%	&	66\%	&	12\%	&	5\%	\\ \hline
r 3	&	100\%	&	100\%	&	54\%	&	11\%	&	4\%	\\ \hline
r 4	&	100\%	&	99\%	&	45\%	&	1\%	&	6\%	\\ \hline
r 5	&	100\%	&	99\%	&	37\%	&	3\%	&	4\%	\\ \hline
r 6	&	100\%	&	99\%	&	30\%	&	12\%	&	12\%	\\ \hline
r 7	&	99\%	&	98\%	&	27\%	&	3\%	&	7\%	\\ \hline
r 8	&	99\%	&	98\%	&	24\%	&	3\%	&	0\%	\\ \hline
r 9	&	99\%	&	97\%	&	19\%	&	3\%	&	5\%	\\ \hline
r 10	&	99\%	&	97\%	&	13\%	&	7\%	&	5\%	\\ \hline
\end{tabular}
\end{center}
Suggest, with reasons, appropriate values for d and the parameter
underlying AR(1) process.
Suggest, with reasons, appropriate values for d and the parameter
underlying AR(1) process.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
3
(i)
0
12
20
Maximum loss: 20
5
0
15
15
%%April 2006
%%Examiners Report
8
3
0
8
Minimax is d 3 .
(ii)
P( 1 ) = 0.15
P( 2 ) = 0.6
P( 3 ) = 0.25
d 1 = 0.15
d 2 = 0.15
d 3 = 0.15
0 + 0.6
5 + 0.6
8 + 0.6
12 + 0.25 20 = 12.2
0 + 0.25 15 = 4.5
3 + 0.25 0 = 3
Hence the Bayes decision is d 3 .
Comments on question 3: No comments given.
4
(i)
k
= Cov(X t , X t k )
= Cov( X t 1 + e t , X t k )
= k 1
and
0
= Cov(X t , X t )
= Cov( X t 1 + e t , X t 1 + e t )
= 2 Cov(X t 1 , X t 1 ) + Cov(e t , e t )
= 2 0 + 2
and hence
2 )
0 (1
=
2
2
i.e.
0 =
1
2
So our solution is
k 2
k =
1
2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% - %%-- Page 3%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% - April 2006
%% - Examiners Report
The autocorrelation function is given by
k
=
k
=
k
0
(ii)
The autocorrelation should decay exponentially as i increases. Looking at the table this behaviour occurs after differencing 2 times, suggesting the value of d = 2.
We know that the ratio of successive r s should be . We can form these
ratios as follows:
r 2 /r 1
r 3 /r 2
r 4 /r 3
r 5 /r 4
r 6 /r 5
r 7 /r 6
r 8 /r 7
r 9 /r 8
r 10 /r 9 80%
82%
83%
82%
81%
90%
89%
79%
68%
Average 81.6%
Alternatively we can take the ith root of the ith autocorrelation:
i ith root
1
2
3
4
5
6
7
8
9
10 83%
81%
81%
82%
82%
82%
83%
84%
83%
82%
Average 82%
Both approaches suggest the value of alpha is around 82%.
Full credit should be given to any reasonable approach.
Comments on question 4: This question was generally done poorly. Although most
candidates made a reasonable attempt at (i), very few correctly identified appropriate values
or sensible reasons in (ii).
\end{document}
