\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}Let X be a random variable with probability density function:
⎧ 1 x
⎪ e
f ( x ) = ⎨ 2
⎪ 1 e − x
⎩ 2
(i)
; x ≤ 0
; x > 0
Show that the moment generating function of X is given by:
M X ( t ) = (1 − t 2 ) − 1 ,
for t < 1 .
(ii)
CT3 A2014–2
[3]
Hence find the mean and the variance of X using the moment generating
function in part (i).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%--- Question 5
\item Consider ten independent random variables X 1 , ... , X 10 which are identically
distributed with an exponential distribution with expectation 4.
10
\begin{enumerate}
\item (i)
Specify the approximate distribution of X = \sum  X i , including all parameters,
i = 1
using the central limit theorem.
6

\item (ii) Calculate the approximate value of the probability P [ X < 40] using the result
in part (i).

\item (iii) Calculate the exact probability P [ X < 40] .
\item (iv) Comment on the answers in parts (ii) and (iii).
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
4
(i)
1
M X ( t )  E ( e ) 
2
tX
1  e ( t  1) x 
 

2   t  1  
0

0
 e
( t  1) x

1  e ( t  1) x 
 

2   t  1  

1
dx   e ( t  1) x dx
2
0

0
and for  t  1
%% --- Page 3 (Probability and Mathematical Statistics) – April 2014 – 
M X ( t ) 
(ii)
1  1
1 
1


 
2  t  1 t  1  1  t 2

M  X ( t )  (1  t 2 )  1   (1  t 2 )  2 (  2 t )  2 t (1  t 2 )  2


 E ( X )  M  X (0)  0

M  X ( t )  2 t (1  t 2 )  2

  2(1  t
2  2
)
 2 t (  2)(1  t 2 )  3 (  2 t )
 2(1  t 2 )  2  8 t 2 (1  t 2 )  3
 E ( X 2 )  M  X (0)  2
V ( X )  E ( X 2 )  E 2 ( X )  2
(Alternatively, based on a series expansion:
M X ( t )  1  t 2  t 4  ...  E ( X )  0 and E ( X 2 )  2 and the variance follows.)

%%-- Generally well answered. In part (ii) most candidates were familiar with the method, but some showed poor differentiation skills.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
5


(i) X ~ N  ,  2 with   10  4  40 and  2  10   4   160
(ii) X is symmetric so P  X  40   0.5
(iii) The exact distribution of X is gamma(10, 1⁄4)
2
 1

P  X  40   P  2   X  20   P  Y  20  where Y has a  2 distribution
4


with 20 d.f.
P  X  40   P  Y  20   0.5421
(iv)
Although the sample size here is small, the CLT gives an answer which is
close to the exact probability.
Mostly well answered. There were a few problems with the distributions in part (iii). In part
(iv) comments should refer to the use of CLT with small samples for full marks.
Page 4 (Probability and Mathematical Statistics) – April 2014 – 
\end{document}
