\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
%%--- Question 2

Compare the characteristics of deterministic and stochastic models, by considering the
relationship between inputs and outputs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
Question 2
A deterministic model is a model which does not contain any random components.
The output is determined once the fixed inputs and the relationships between inputs and
outputs have been defined.
A stochastic model is one that recognises the random nature of the input components.
The inputs to a stochastic model are random variables, and hence for any given values of the
inputs the outputs are an estimate of the characteristics of the model.
Several independent iterations of the model are required for each set of inputs to study their
implications.
The output of a stochastic model gives the distribution of relevant results for a distribution
of scenarios.
A deterministic model can be seen as a special case of a stochastic model.
The output of a stochastic model can be reproduced if the same random seed is used.
The output of a deterministic model is only a snap shot or an estimate of the characteristics
of the model for a given set of inputs.
Full marks could be obtained for rather less than is written above. The maximum number
of marks attainable was 4 even if all the above points were made.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
%%--- Question 2

Compare the characteristics of deterministic and stochastic models, by considering the
relationship between inputs and outputs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
%%--- Question 3
3 The government of a small island state intends to set up a model to analyse the
mortality of the island’s population over the past 50 years.
Describe the process that would be followed to carry out the analysis.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55

\end{document}
