\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}


[Total 9]10
11
A study is undertaken of marriage patterns for women in a country where bigamy is
not permitted. A sample of women is interviewed and asked about the start and end
dates of all their marriages and where the marriages had ended, whether this was due
to death or divorce (all other reasons can be ignored). The investigators are interested
in estimating the rate of first marriage for all women and the rate of re-marriage
among widows.
\begin{enumerate}
\item (i) Draw a diagram illustrating a multiple-state model which the investigators
could use to make their estimates, using the four states: “Never married”,
“Married”, “Widowed” and “Divorced”.

\item (ii) Derive from first principles the Kolmogorov differential equation for first
marriages.
\item 
(iii) Write down the likelihood of the data in terms of the waiting times in each
state, the numbers of transitions of each type, and the transition intensities,
assuming the transition intensities are constant.
\item 
(iv) Derive the maximum likelihood estimator of the rate of first marriage.
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Question 10
(i)
(ii)
1 Never married 2 Married
3 Widowed 4 Divorced
Using the numbering of the states above, let the probability that a women who is in
state i at time x will be in state j at time x + t be t p ij x .
Using the Markov property, and conditioning on the state occupied
at time x + t,
and noting that for first marriages return from the widowed or
divorced state is not possible, we can write
t + dt
11
12
12
22
p 12
x = t p x dt p x + t + t p x dt p x + t
Using the law of total probability,
dt
p x 22 + t = 1 − dt p x 23 + t − dt p x 24 + t ,
so that
t + dt
11
12
12
23
24
p 12
x = t p x dt p x + t + t p x (1 − dt p x + t − dt p x + t )
Let the transition rate from state i to state j at time x+t be \mu ijx + t .
Assume that
Page 12
dt
p ij x + t = \mu ij x + t dt + o ( dt ) , i ≠ j %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% — 
o ( dt )
= 0 .
dt → 0 + dt
where lim
Substituting for the
t + dt
t
p ij x in the equation above produces
11 12
12
23
24
p 12
x = t p x \mu x + t dt + t p x (1 − \mu x + t dt − \mu x + t dt ) + o ( dt )
Therefore
t + dt
12
11 12
12 23
12 24
p 12
x − t p x = t p x \mu x + t dt − t p x \mu x + t dt − t p x \mu x + t dt + o ( dt )
and, taking limits, we have
lim
dt → 0 +
t + dt
12
p 12
12
12 23
12 24
x − t p x
= t p 11
x \mu x + t − t p x \mu x + t − t p x \mu x + t
dt
So
d 12
11
12
12 23
12 24
t p x = t p x \mu x + t − t p x \mu x + t − t p x \mu x + t
dt
%%%%%%%%%%%%%%%%%%
\newpage
\begin{itemize}
\item (iii)
Let the waiting time in state i be v i ,
and the number of transitions from state i to state j be d ij
and the transition intensity from state i to state j is \mu ij
\item Then the likelihood, L, may be written
\[L = K exp[ − v 1 \mu 12 − v 2 ( \mu 23 + \mu 24 ) − v 3 \mu 32 − v 4 \mu 42 ] \mu 12 d 12 \mu 23 d 23 \mu 24 d 24 \mu 32 d 32 \mu 42 d 42 .\]
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%
(iv)
The logarithm of the likelihood is
log e L = log e K − v 1 \mu 12 − v 2 ( \mu 23 + \mu 24 ) − v 3 \mu 32 − v 4 \mu 42
+ d 12 log e \mu 12 + d 23 log e \mu 23 + d 24 log e \mu 24 + d 32 log e \mu 32 + d 42 log e \mu 42
Differentiating with respect to \mu 12 gives
d
\partial L
= − v 1 + 12 .
\partial\mu 12
\mu 12
Page 13 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% — 
Setting this equal to 0 and solving for \mu 12 gives
^
\mu 12 =
d 12
.
v 1
This is a maximum because
\partial 2 L
\partial\mu 12
2
=−
d 12
\mu 12 2
which is negative.
Answers to (i), (iii) and (iv) were generally good. In (i) the arrows from Widowed to Married
and Divorced to Married were not required for full marks, as the question is about first
marriages. Answers to (ii) were more disappointing, with many candidates omitting steps in
32
14 42
the argument. In (ii), some candidates included the extra terms + t p 13
x \mu x + t + t p x \mu x + t . Since it
is just about possible to interpret the question in a way such that these should be included,
this was not heavily penalised.
\end{itemize}
\end{document}
