\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}


\large
%%- Question 4 
\noindent A manufacturing company is analysing its accident record. The accidents fall into two categories:


\begin{itemize}
    \item \textbf{Minor} : dealt with by first aider. Average cost \$50.
    \item \textbf{Major} : hospital visit required. Average cost \$1,000.
\end{itemize}  
The company has 1,000 employees, of which 180 are office staff and the rest work in the factory.\\

\noindent The analysis shows that 10\% of employees have an accident each year and 20\% of  accidents are major. It is assumed that an employee has no more than one accident in
a year.
\begin{enumerate}[(a)]
\item  Determine the expected total cost of accidents in a year. \\
\medskip 
On further analysis it is discovered that a member of office staff has half the probability of having an accident relative to those in the factory.
\item  Show that the probability that a given member of office staff has an accident in a year is 0.0549. 
\item  Determine the probability that a randomly chosen employee who has had an accident is office staff. 
\end{enumerate}

\newpage
%% Q4
%% (i)

\noindent \textbf{Notation}
\begin{itemize}
\item A = Accident
\item O = Office Staff 
\item F = Factory Staff.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%
\medskip 
\noindent \textbf{Part (a) }\begin{eqnarray*}
\mbox{ Expected cost of an accident } &=& \mbox{Average Cost(Major) } \times P(\mbox{Major}|A)
+  \\ 
& &  \mbox{Average Cost(Minor)}  \times P(\mbox{Minor}|A)\\
&=& 0.2\times 1000 \;+\; 0.8 \times 50 \\
&=& \$240 \\ 
\end{eqnarray*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray*}
\mbox{ Expected total cost of accidents } &=&
\$ 240 \times 1000 \times  0.1\\
&=& \$24,000 \\
\end{eqnarray*}
\newpage
\noindent \textbf{Part (b) }\\
We are told that \[P(A|O) = 0.5 P(A|F)\] or equivalently \[2 P(A|O) = P(A|F).\]
Also we know that 
\begin{itemize}
    \item $P(A) \;=\; 0.1$
    \item $P(O) \;=\; 0.18$
    \item $P(F) \;=\; 0.82$
\end{itemize}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray*}
P(A) &=& P(A \mbox{ and } O)+P(A \mbox{ and }F)\\
& & \\
&=& P(A|O)P(O)+P(A|F)P(F) \\ & & \\ &=& P(A|O)\times 0.18 + 2\times P(A|O)\times 0.82 \\
& & \\
0.1 &=& P(A|O) \times 1.82
\end{eqnarray*}
\[P(A|O) \;=\; \frac{0.1 }{1.82} \;=\;  0.0549\]

%%%%%%%%%%%%%%%%%%%%%%%%%%% 

\noindent \textbf{Part (c) }
\begin{eqnarray*}
P(O|A) &=& \frac{P ( A | O ) P ( O )}{P ( A )}\\ 
& &  \\
&=& \frac{0.0549\times 0.18}{0.1}\\
& &  \\
&=&  0.099.\\
\end{eqnarray*}
% [2]
% [TOTAL 7]
% Part (i) was well answered with most candidates giving a fully or partially correct answer. Answers in parts (ii) and (iii) were generally poor; some candidates were confused with the conditional probabilities and considered conditioning on the wrong event.

\end{document}
