\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
7
[3]
For a group of policies the total number of claims arising in a year is modelled as a
Poisson variable with mean 10. Each claim amount, in units of £100, is independently
modelled as a gamma variable with parameters = 4 and = 1/5.
Calculate the mean and standard deviation of the total claim amount.
CT3 A2005
2
[5]8
9
The distribution of claim size under a certain class of policy is modelled as a normal
random variable, and previous years records indicate that the standard deviation is
£120.
(i) Calculate the width of a 95% confidence interval for the mean claim size if a
sample of size 100 is available.
[2]
(ii) Determine the minimum sample size required to ensure that a 95% confidence
interval for the mean claim size is of width at most £10 .
[2]
(iii) Comment briefly on the comparison of the confidence intervals in (i) and (ii)
with respect to widths and sample sizes used.
[1]
[Total 5]
Let X 1 , , X n denote a large random sample from a distribution with unknown
population mean and known standard deviation 3. The null hypothesis H 0 : = 1 is
to be tested against the alternative hypothesis H 1 : > 1, using a test based on the
sample mean with a critical region of the form X
k , for a constant k.
It is required that the probability of rejecting H 0 when = 0.8 should be
approximately 0.05, and the probability of not rejecting H 0 when = 1.2 should be
approximately 0.1.
(i)
Show that the test requires
k 0.8
3/ n
where
(ii)
0.95 and
k 1.2
3/ n
0.10
is the standard normal distribution function.
[4]
The values for the sample size n and the critical value k which satisfy the
requirements of part (i) are n = 482 and k = 1.025 (you are not asked to verify
these values).
Calculate the approximate level of significance of the test, and comment on
the value.
[3]
[Total 7]
CT3 A2005
3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let S = X 1 + X 2 + . . . + X N be the total claim amount.
Note that E[N] = V[N] = 10, E[X] = 4/(1/5) = 20, V[X] = 4/(1/5) 2 = 100
E[S] = E[N]E[X] = 10(20) = 200
mean of total claim amount = £20,000.
V[S] = E[N]V[X] + V[N]{E[X]} 2
=10(100) + 10(20) 2 = 5000
s.d.[S] = 70.71
8
(i)
s.d. of the total claim amount = £7,071
Width of 95% confidence interval: 1.96
120
120
[or 2 1.96
n
n
3.92
120
]
n
120
23.52
100
120
[or 2 1.96
£47.04]
100
1.96
(ii)
For the width of a 95% confidence interval to be at most
120
1.96
10
n
n
1.96 120
10
23.52 , n
10 we require
553.19
i.e., take the sample size as 554.
Page 3Subject CT3 (Probability and Mathematical Statistics Core Technical)
9
(iii) The confidence interval in (ii) in narrower
much larger sample size.
(i) X approx
0.05
N
,
3 2
n
April 2005
Examiners Report
to achieve this we require a
for large n by the central limit theorem.
P(reject H 0 | = 0.8) = P( X > k| = 0.8)
k 0.8
3/ n
= 1
k 0.8
3/ n
0.95
and
0.1
P(do not reject H 0 | = 1.2)
= P ( X
1.2)
k 1.2
3/ n
=
(ii)
k |
Significance level
1
= P(reject H 0 when H 0 is true) = P ( X
1.025 1
3 / 482
1
k |
1)
(0.18) 1 0.57 0.43.
The significance level of the test is very high (43%).
