The number of claims made by each policyholder in a certain class of business is
modelled as having a Poisson distribution with mean λ.
(i)
Derive an expression for the probability, p, that a policyholder in this class has
made at least one claim.
[2]
The claims records of 20 randomly chosen policyholders were examined and the
number of policyholders that made at least one claim in a year, X, was recorded.
(ii)
(a) State the distribution of the random variable X and its parameters.
(b) Derive an expression for the maximum likelihood estimator of the
probability p given in (i) using your answer in (ii)(a).
[4]
(iii)
Show that, in the case X = 5, the maximum likelihood estimate (MLE) of p is
[3]
p ˆ = 0.25 and hence calculate the MLE of λ.
It is now found that of the five policyholders who had made at least one claim there
were four who had made exactly one claim and one who had made two claims.
(iv)
Calculate the MLE of λ given this additional information.
CT3 S2011—3
[4]
[Total 13]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


6
(i) p = Pr( N ≥ 1) = 1 − Pr( N = 0) = 1 − e −λ .
(ii) (a)
X ~ Bin(20,p)
(b)
L ( p ) ∝ p X (1 − p ) 20 − X
⇒ l ( p ) = log( L ( p )) = X log p + (20 − X ) log(1 − p )
and l '( p ) = 0 ⇒
(and l ''( p ) = −
(iii)
p ˆ =
X 20 − X
X
−
= 0 ⇒ X − 20 p ˆ = 0 ⇒ p ˆ =
p ˆ
1 − p ˆ
20
20 − X
X
−
≤ 0)
p ˆ 2 (1 − p ˆ ) 2
5
= 0.25
20
Then, using the invariance property of the MLE:
ˆ
p ˆ = 1 − e −λ ⇒ λ ˆ = − log(1 − p ˆ ) = − log(0.75) = 0.288
Likelihood function now is:
(iv)
L ( λ ) ∝ P ( X = 0) 15 × P ( X = 1) 4 × P ( X = 2)
( ) ( λ e ) ( λ e )
∝ e −λ
15
−λ
4
2 −λ
⇒ l ( λ ) ∝ − 15 λ + 4 log λ − 4 λ + 2 log λ − λ = − 20 λ + 6 log λ
and l ′ ( λ ) = 0 ⇒ − 20 +
Page 4
6
= 0 ⇒ λ ˆ = 0.3
ˆ λSubject CT3 (Probability and Mathematical Statistics) — September 2011 — Examiners’ Report
(Also l ′′ ( λ ) = −
6
< 0 , hence max)
λ ˆ 2
Notice that part (ii)(b) requires the use of the binomial distribution from (ii)(a). In part (iii)
the invariance property must be used and mentioned for full credit.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
7
(i) F = 0.213/0.0106 = 20.094 and at the 5% significance level,
F 3,20 (0.05) = 3.098.
Since F = 20.094 > 3.098, there is strong evidence against the null hypothesis,
and we conclude that there are differences in the mean amounts paid out by
the companies.
(ii) The variance of the residuals seems to be similar for the four companies; this
is consistent with the assumption of constant variance in the response variable.
Also there are no obvious patterns or outliers. The analysis seems valid.
(iii) (a)
LSD = t 20 (0.025)
= 2.086
(b)
⎛ 1 1 ⎞
σ 2 ⎜ + ⎟
⎝ 6 6 ⎠
0.0106 / 3 = 0.124
The four company (treatment) means are:
y 1 i = 17.810
18.940
17.715
= 2.968, y 2 i =
= 3.157, y 3 i =
= 2.953
6
6
6
y 4 i = 16.185
= 2.698
6
which are given in order and underlined as
y 4 i < y 3 i < y 1 i < y 2 i
Amounts paid out by companies 2 and 4 are significantly different
from those paid out by the other two companies. Company 4 seems to
pay out significantly lower amounts, with Company 2 paying
significantly higher.
Parts (i) and (ii) were generally well answered. In part (iii) many candidates did not use the
correct formula for LSD and then performed pair-wise comparisons using the wrong statistic.
Page 5Subject CT3 (Probability and Mathematical Statistics) — September 2011 — Examiners’ Report
