4
The random variables X and Y are related as follows:
X conditional on Y = y has a N(2y, y 2 ) distribution.
Y has a N(200, 100) distribution.
Derive the unconditional variance of X, V[X].
CT3 S2011—2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%555
[3]5
Consider the random variable X taking the value X = 1 if a randomly selected person
is a smoker, or X = 0 otherwise. The random variable Y describes the amount of
physical exercise per week for this randomly selected person. It can take the values 0
(less than one hour of exercise per week), 1 (one to two hours) and 2 (more than two
hours of exercise per week). The random variable R = (3 − Y) 2 (X + 1) is used as a risk
index for a particular heart disease.
The joint distribution of X and Y is given by the joint probability function in the
following table.
X
0
1
6
0
0.2
0.1
Y
1
0.3
0.1
2
0.25
0.05
(i) Calculate the probability that a randomly selected person does more than two
hours of exercise per week.
[1]
(ii) Decide whether X and Y are independent or not and justify your answer. [2]
(iii) Derive the probability function of R. [3]
(iv) Calculate the expectation of R.
[2]
[Total 8]


4
V[X]
= E[V(X|Y)] + V[E(X|Y)]
= E[Y 2 ] + V[2Y] = (100 + 200 2 ) + 4*100 = 40,500
Generally very well answered.
5
(i) P[Y = 2] = 0.25 + 0.05 = 0.3
(ii) P[X = 0] = 0.75 and
P[{X = 0} ∩ {Y = 2}] = 0.25 ≠ 0.225 = 0.3 * 0.75 = P[X = 0] * P[Y = 2]
Therefore X and Y are not independent.
(Any other joint probability can be used.)
Page 3Subject CT3 (Probability and Mathematical Statistics) — September 2011 — Examiners’ Report
(iii)
(iv)
The probability function is
r 1
2
4
8
9
18
P(R = r) 0.25 0.05 0.3 0.1 0.2 0.1
E[R] = 0.2 * 9 + 0.3 * 4 + 0.25 * 1 + 0.1 * 18 + 0.1 * 8 + 0.05 * 2
= 1.8 + 1.2 + 0.25 + 1.8 + 0.8 + 0.1 = 5.95
In part (ii) notice that one example of P(XY) = P(X)P(Y) is not sufficient for showing
independence (it needs to hold for all cases). Also, some candidates failed to provide the
probability function in (iii).

