\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

\begin{enumerate}

[Total 12]
9 A survey team is using satellite technology to measure the height of a mountain. This
is an established technology and the variability of measurements is known. On each
satellite pass over a mountain they get a measurement that they know lies within ±5m
of the true height with a 95% probability. The survey height is given by the mean of
the measurements. They assume that all measurements are independent and follow a
normal distribution with mean equal to the true height.
(i) (a) Show that the standard deviation of a single measurement is 2.551m.
(b) Determine how many satellite passes over a mountain are required to
have a 95% confidence interval for the true height with width less than
1m.
[5]
In a mountain range there are two summits which appear to have a similar height.
The team manages to get 20 measurements for each summit and finds there is a
difference of 1.6m between the mean survey height of the two summits.
(ii) Perform a statistical test of the null hypothesis that the summits’ true heights
are the same, against the alternative that they are different. [4]
At the same time the team is testing a new system on these two summits. They again
get 20 measurements on each summit with an estimated standard deviation on the first
summit of 2.5m and on the second of 2.6m. This system also measures the difference
in survey heights between the two peaks to be 1.6m.
(iii) Perform a statistical test of the same hypotheses as in part (ii) when heights
are measured by the new system, justifying any assumptions you make. [5]
(iv) Comment on your answers to parts (ii) and (iii). [3]
[Total 17]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

