
\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

\item The following 24 observations give the length of time (in hours, ordered) for which a
specific fully charged laptop computer will operate on battery before requiring
recharging.
\begin{verbatim}
1.2 1.4 1.5 1.6 1.7 1.7 1.8 1.8 1.9 1.9 2.0 2.0
2.1 2.1 2.1 2.2 2.3 2.4 2.4 2.5 3.1 3.6 3.7 4.5
\end{verbatim}

The owner of this computer is about to watch a film on his fully charged laptop.
Calculate from these data the longest showing time for a film that he can watch, so
that the probability that the battery's lifetime will be sufficient for watching the entire
film is 0.75.

\end{enumerate}
\newpage
%%--- 6
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Page 2%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5 – April 2012 – %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
1
We want the first quartile of the data.
⎛ n + 2 ⎞
Q 1 = ⎜
⎟ th observation counting from below = 6.5th observation
⎝ 4 ⎠
=
1.7 + 1.8
= 1.75 hours.
2
[With alternative definition:
⎛ n + 1 ⎞
Q 1 = ⎜
⎟ th observation counting from below = 1.725 ]
⎝ 4 ⎠
Most answers were quite poor. Many candidates tried to work with a normal or t distribution,
when this was not justified (or required). Only a small number of candidates realised that
quartiles were required – but then a large proportion of them used the wrong quartile.
\end{document}
