\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

[Total 8]
%%--- Question 7
The heights of adult males in a certain population are Normally distributed with
unknown mean \mu and standard deviation \sigma = 15.
Prior beliefs about \mu are described by a Normal distribution with mean 187 and
standard deviation 10.
(i)
Calculate the prior probability that \mu is greater than 180.

A sample of 80 men is taken and the mean height is found to be 182.
8
(ii) Calculate the posterior probability that \mu is greater than 180.

(iii) Comment on your results from parts (i) and (ii).
(i) (a) Write down the Box-Muller algorithm for generating samples from a
standard Normal distribution.
(b) Give an advantage and a disadvantage of the Box-Muller algorithm
relative to the Polar method.

[Total 8]

(ii)
Extend the algorithm in part (i) to generate samples from a Lognormal
distribution with parameters \mu and \sigma 2 .

A portfolio of insurance policies contains n independent policies. The probability of a
claim on a policy in a given year is p and the probability of more than one claim is
zero. Claim amounts follow a Lognormal distribution with parameters \mu and \sigma 2 . The
insurance company is interested in estimating the probability \theta that aggregate claims
exceed a certain fixed level M.
(iii)
Construct an algorithm to simulate aggregate annual claims from this
portfolio.

The insurance company estimates that \theta is around 10%.
(iv)
Calculate the smallest number of simulations the insurance company should
undertake to be able to estimate \theta to within 1% with 95% confidence.

The insurance company is considering the impact on \theta of entering into a reinsurance
arrangement.
(v)
CT6 A2014–4
Explain whether the insurance company should use the same pseudo random
numbers when simulating the impact of reinsurance.


7
(i)
P(\mu > 180) = P(N(187, 10 2 ) > 180)
180 − 187 ⎞
⎛
= P ⎜ N (0,1) >
⎟
10
⎝
⎠
= P(N(0,1) > − 0.7)
= 0.75804
(ii)
We know that \mu|x ~ N ( \mu * , \sigma * 2 )
1 ⎞
⎛ 80 × 182 187 ⎞ ⎛ 80
Where \mu * = ⎜
+ 2 ⎟ ⎜ 2 + 2 ⎟ = 182.14
2
10 ⎠ ⎝ 15 10 ⎠
⎝ 15
And \sigma * 2 =
so
1
80
1
+ 2
2
15 10
= 2.73556 = 1.6540 2
P (\mu > 180) = P ( N (182.14, 1.654 2 ) > 180)
180 − 182.14 ⎞
⎛
= P ⎜ N (0,1) >
⎟
1.654
⎝
⎠
= P ( N (0,1) > -1.29192)
= 0.38 × 0.9032 + 0.62 × 0.90147
= 0.90180
(iii)
The probability has risen, reflecting our much greater certainty over the value
of \mu as a result of taking a large sample.
Page 7%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5 – April 2014 – 
This is despite the fact that our mean belief about \mu has fallen, which a priori
might make a lower value of \mu more likely.
The posterior distribution has thinner tails / lower volatility, since we have
increased credibility around the mean
This question was mostly well answered. A small number of candidates were not aware that
the formulae for the Normal / Normal model are given in the tables, and therefore struggled
with the algebra required to derive the posterior distribution.
8
(i)
(a)
Let u 1 and u 2 be independent samples from a U (0,1) distribution.
Then Z 1 = − 2 log u 1 cos(2 \pi u 2 )
Z 2 = − 2 log u 1 sin(2 \pi u 2 )
are independent standard normal variables.
(b)
Advantage –
generates a sample of every pair of u 1 and u 2 – no
possibility of rejection.
Disadvantage – requires calculation of sin and cos functions which is
more computationally intensive.
(ii)
Generate Z as in (i). Then
Y = exp(\mu + \sigma Z )
is a sample from the required Lognormal distribution.
(iii)
Set
X = 0, k = 0
Step 1 generate a sample u from U (0,1), set k = k + 1
Step 2 If u ≤ p then go to step 3 else go to step 4
Step 3 Generate a sample Y from the Lognormal distribution in (ii) and set
X = X + Y
Step 4 If k = n finish else go to step 1
X represents aggregate claims on the portfolio.
Page 8%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5 – April 2014 – 
(iv)
\theta ˆ (1 − \theta ˆ )
0.09
=
.
n
n
The standard error will be approximately
We want
i.e.
n >
0.09
× 1.96 < 0.01
n
0.09 × 1.96
= 58.8
0.01
i.e. n > 3457.44
so 3,458 simulations are needed.
(v)
The insurer should use the same pseudo-random numbers so that any variation
in simulation results is due to the impact of the reinsurance and not just due to
random variation in the simulation process.
Parts (i) and (v) were well answered. The remaining parts were found by many candidates to
be the hardest questions on the paper. In part (ii) many candidates could not relate the
Lognormal distribution to the Normal distribution from which samples had been generated in
(i). Only the best candidates attempted parts (iii) and (iv).
