\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

[Total 9]4
As part of a simulation study an actuary is asked to design an algorithm for simulating
claims from a particular type of insurance policy. The probability distribution of the
annual number of claims on a policy is given by:
Probability
No claims One claim Two claims
0.4 0.4 0.2
The claim size distributions of the first and second claims are different. The size of
the first claim follows an exponential distribution with mean 10. The size of the
second claim follows a Weibull distribution with parameters c = 1 and γ = 4.
5
(i) Construct an algorithm to simulate the first claim on a given policy. 
(ii) Construct an algorithm to simulate the second claim on a given policy. 
(iii) Construct an algorithm to simulate the total annual claims on a given policy.

[Total 10]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The table below shows the incremental claims incurred for a certain portfolio of
insurance policies.
Accident year
2011
2012
2013
Development year
0
1
2
2,233
3,380
4,996
1,389
1,808
600
Cumulative numbers of claims are shown in the following table:
Accident year
2011
2012
2013
Development year
0
1
2
140
180
256
203
230
224
(i) Calculate the outstanding claim reserve for this portfolio using the average
cost per claim method with grossing up factors.
[7]
(ii) State the assumptions underlying the calculations in part (i).
CT6 S2014–3

[Total 10]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

4
(i)
Using the inversion method, set
u = F(x) = 1  e


x
10
x
10
i.e. 1  u = e
i.e. x
log(1  u) = 10
i.e. x = 10 log(1  u)
so the algorithm is:
Step 1
Step 2
Generate a sample u from a U(0,1) distribution.
Set x = 10 log(1  u).

Page 5%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% – September 2014 – Examiners’ Report
(ii)
Again using the inversion method, set
 x
u = F(x) = 1  e
i.e.  x
1  u = e
i.e. x 4 = log(1  u)
i.e. x = [log(1  u)] 1⁄4
4
4
so the algorithm is
Step 1
Step 2
Generate a sample u from a U(0,1) distribution.
Set x = [log(1  u)] 1⁄4 .

(iii)
Our algorithm is as follows:
Step 1 Generate a sample u from a U(0,1) distribution.
Step 2 If 0  u  0.4 then total claim amount X = 0 and stop
Else continue to step 3.
Step 3 If 0.4 < u 0.8 then simulate a claim from Exp(1/10)
distribution using the algorithm in (i) and set X = this value
and stop.
Else go to step 4.
Step 4
Simulate claims using the algorithms in (i) and (ii) and set
X = total of the two simulated claims.

[Total 10]
This question was well answered.
\end{document}
