\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
3
[6]
Andy is a famous weight lifter who will be competing at the Olympic Games. He has
taken out special insurance which pays out if he is injured. If the injury is so serious
that his career is ended the policy pays \$1m and is terminated. If he is injured but
recovers the insurance payment is \$0.1m and the policy continues.
The insurance company’s underwriters believe that the probability of an injury in any
year is 0.2, and that the probability of more than one injury in a year can be ignored.
If Andy is injured, there is a 75\% chance that he will recover.
Annual premiums are paid in advance, and the insurance company pays claims at the
end of the year. Assume that this is the only policy that the insurance company
writes, and that it has an initial surplus of \$0.1m.

\begin{enumerate}
    \item (i) Define what is meant by $\psi(\$0.1m,1)$ and \psi(\$0.1m).
\item (ii) Calculate the annual premium charged assuming the insurance company uses a
premium loading of 30\%.
\item 
(iii) Determine $\psi(\$0.1m, 2)$.
\end{enumerate}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
3
(i)
If U  t   U  ct  S  t  where U  U  0   $0.1 m then
\psi $0.1 m ,1   Pr  U  t   0 for some t   0,1  given U  0   $0.1 m 
and
\psi $0.1 m   Pr( U  t   0 for some t  0 given U  0   $0.1 m )
(ii)
The premium charged will be:
1.3  0.2  (0.25  $1m  0.75  $0.1m)  $0.0845m.

(iii)
The possibilities are tabulated below, where N means not injured, R means
injured but recovered and X means injured but career ending:
Year 1 Year 2 Probability Ruin?
N N 0.8  0.8 =0.64 No
N R 0.8  0.15 = 0.12 No
N X 0.8  0.05 = 0.04 Yes
R N 0.15  0.8 = 0.12 No
R R 0.15  0.15 = 0.0225 No
R X 0.15  0.05 = 0.0075 Yes
X N/A 0.05 Yes
Summing the cases where ruin occurs we have:
 ($0.1m, 2)  0.04  0.0075  0.05  0.0975

% Many candidates lost marks in part (i) by not giving a sufficiently precise definition to score full marks. For part (iii) candidates who 
% worked through the possibilities methodically generally scored well. A number of candidates unnecessarily used approximate methods in
% part (iii).

\end{document}

