\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
5
An investigation into the mortality experience of a sample of the male student population of a large university has been carried out. The university authorities wish to know whether the mortality of male students at the university is the same as that of
males in the country as a whole. They have drawn up the following table.
\begin{verbatim}
Age x Number of deaths Expected number
of deaths assuming
national mortality
18
19
20
21
22
23 13
15
14
20
12
8 10
12
14
12
8
5
\end{verbatim}
Carry out an overall test of the university authorities’ hypothesis, stating your
conclusion.
CT4 S2008—3

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\begin{itemize}
\item Using the chi-squared test (a suitable overall test).
actual deaths - expected deaths
, then the test statistic is
expected deaths
where m is the number of ages, which in this case is 6.
If z x =
\sum z x 2 ∼ \chi^2_m ,
x
\item The calculations are shown below.
\begin{verbatim}
Age x z x z x 2
18
19
20
21
22
23 0.9487
0.8660
0
2.3094
1.4142
1.3416 0.9
0.75
0
5.3333
2
1.8
\end{verbatim}
\item 
Therefore the value of the test statistic is 10.783.
\item The critical value of the chi-squared distribution
at the 5% level of significance with 6 degrees of
freedom is 12.59.
\item Since 10.783 < 12.59 there is insufficient evidence to reject
the hypothesis that the mortality rate of men in the University is the same as that of
the national population.
\end{itemize}
%%Page 6Subject CT4 — Models Core Technical — September 2008 — Examiners’ Report
\end{document}
