\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}
\item %%- Question 8
An education authority provides children with musical instrument tuition. The
authority is concerned about the number of children giving up playing their
instrument and is testing a new tuition method with a proportion of the children which
it hopes will improve persistency rates. Data have been collected and a Cox
proportional hazards model has been fitted for the hazard of giving up playing the
instrument. Symmetric 95\% confidence intervals (based upon standard errors) for the
regression parameters are shown below.
Covariate
Confidence Interval
Instrument
Piano
Violin
Trumpet 0
[-0.05,0.19]
[0.07,0.21]
Tuition method
Traditional
New 0
[-0.15,0.05]
Sex
Male
Female
[-0.08,0.12]
0

\begin{enumerate}[(i)]
\item (i) Write down a general expression for the Cox proportional hazards model,
defining all terms that you use.

\item (ii) State the regression parameters for the fitted model. 
\item (iii) Describe the class of children to which the baseline hazard applies. [1]
\item (iv) Discuss the suggestion that the new tuition method has improved the chances
of children continuing to play their instrument.
\item 
(v) Calculate, using the results from the model, the probability that a boy will still
be playing the piano after 4 years if provided with the new tuition method,
given that the probability that a girl will still be playing the trumpet after 4
years following the traditional method is 0.7.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

8
\begin{itemize}
\item (i)

\[h ( z , t ) = h 0 ( t ).exp( \beta . z i T )\]
where h(z,t) is the hazard at duration t
h o (t) is the baseline hazard
z i are the covariates
\beta is the vector of regression parameters
\item (ii)
(iii)
z 1 = 1 plays violin, 0 otherwise \beta 1 = 0.07
z 2 = 1 plays trumpet, 0 otherwise \beta 2 = 0.14
z 3 = 1 new tuition method, 0 otherwise \beta 3 = − 0.05
z 4 = 1 male, 0 otherwise \beta 4 = 0.02
\item Baseline hazard refers to
a female,
following traditional tuition method,
playing the piano
(iv)
\item The parameter associated with the new tuition
method is - 0.05. Because the parameter is negative, the hazard of dropping
out is reduced by the new tuition method.
\item Therefore the new tuition method does appear
to improve the chances of a child continuing
with his or her instrument.
\item However the 95\% confidence interval for the parameter spans zero.
\item So at the 5% significance level it is not possible to conclude that the new
tuition method has improved the chances of children continuing to play their
instrument.
\item (v)
The hazard for a girl being taught the trumpet by the traditional
method giving up is h 0 ( t ) exp(0.14) .
\item Therefore the probability of her still playing after 4 years is
4
⎛ 4
⎞
⎛
⎞
S female (4) = exp ⎜ − \int h 0 ( t ) exp(0.14) dt ⎟ = exp ⎜ − 1.150274 \int h 0 ( t ) dt ⎟
0
⎝ 0
⎠
⎝
⎠
Page 13%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5 — April 2008 — Examiners’ Report
\item Since this is equal to 0.7, we have
4
⎛
⎞
exp ⎜ − 1.150274 \int h 0 ( t ) dt ⎟ = 0.7, so that
0
⎝
⎠
4
log e 0.7 = − 1.150274 \int h 0 ( t ) dt ,
0
4
\end{itemize}

and hence
\int h 0 ( t ) dt =
0
log e 0.7
= 0.310078.
− 1.150274
The hazard of giving up for a boy taught the piano by the new
method is \[h 0 ( t ) exp( − 0.05 + 0.02) = h 0 ( t ) exp( − 0.03).\]
Therefore the probability of him still playing after 4 years is
⎛ 4
⎞
S male (4) = exp ⎜ − \int h 0 ( t ) exp( − 0.03) dt ⎟ = exp [ − 0.310078(0.970446) ]
⎝ 0
⎠
which is exp( - 0.300914) = 0.74014.
\newpage
ALTERNATIVELY
The hazard of giving up for a girl being taught the trumpet by the
traditional method is $h 0 ( t ) exp( \beta 2 )$ .
Therefore the probability of her still playing after 4 years is
4
⎛ 4
⎞
⎛
⎞
S female (4) = exp ⎜ − \int h 0 ( t ) exp( \beta 2 ) dt ⎟ = exp ⎜ − exp( \beta 2 ) \int h 0 ( t ) dt ⎟
0
⎝ 0
⎠
⎝
⎠
and hence
4
\int h 0 ( t ) dt =
0
log e [ S female (4)]
− exp \beta 2
= − exp( − \beta 2 ) log e [ S female (4)] .
The hazard of a boy being taught the piano by the new
method giving up is h 0 ( t ) exp( \beta 3 + \beta 4 ) .
Page 14%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5 — April 2008 — Examiners’ Report
Therefore the probability of him still playing after 4 years is
4
⎛
⎞
S male (4) = exp ⎜ − exp( \beta 3 + \beta 4 ) \int h 0 ( t ) dt ⎟ .
0
⎝
⎠
4
Substituting for
\int h ( t ) dt produces
0
0
S male (4) = exp ( exp( \beta 3 + \beta 4 ) exp( − \beta 2 ) log e [ S female (4)] )
= exp[exp(-0.05+0.02)exp(-0.14)log e (0.7)]
= exp[0.970446 x 0.869358 x -0.356675)
= 0.74014.
