\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

A charity issues a large number of certificates each costing £10 and each being
repayable one year after issue. Of these certificates, 1% are randomly selected to
receive a prize of £10 such that they are repaid as £20. The remaining 99% are repaid
at their face value of £10.
\begin{enumerate}[(i)]
\item (i)
Show that the mean and standard deviation of the sum repaid for a single
purchased certificate are £10.1 and £0.995 respectively.

Consider a person who purchases 200 of these certificates.
\item (ii)
Calculate approximately the probability that this person is repaid more than
£2,040 by using the Central Limit Theorem applied to the total sum repaid.

\item (iii)
An alternative approach to approximating the probability in (ii) above is based
on the number of prize certificates the person is found to hold. This number
will follow a binomial distribution.
Use a Poisson approximation to this binomial distribution to approximate the
probability that this person is repaid more than £2,040.

\item (iv)
Comment briefly on the comparison of the two approximations above given
that the exact probability using the binomial distribution is 0.0517.
\end{enumerate}

\end{document}
