\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}




6 (i) Define the stationary distribution of a Markov chain. 
A baseball stadium hosts a match each evening. As matches take place in the
evening, floodlights are needed. The floodlights have a tendency to break down. If
the floodlights break down, the game has to be abandoned and this costs the stadium
$10,000. If the floodlights work throughout one match there is a 5% chance that they
will fail and lead to the abandonment of the next match.
The stadium has an arrangement with the Floodwatch repair company who are
brought in the morning after a floodlight breakdown and charge $1,000 per day. There
is a 60% chance they are able to repair the floodlights such that the evening game can
take place and be completed without needing to be abandoned. If they are still broken
the repair company is used (and paid) again each day until the lights are fixed, with
the same 60% chance of fixing the lights each day.
(ii) Write down the transition matrix for the process which describes whether the
floodlights are working or not. 
(iii) Derive the long run proportion of games which have to be abandoned. [3]
The stadium manager is unhappy with the number of games being abandoned, and
contacts the Light Fantastic repair company who are estimated to have an 80% chance
of repairing floodlights each day. However Light Fantastic will charge more than
Floodwatch.

(iv) Calculate the maximum amount the stadium should be prepared to pay Light
Fantastic to improve profitability. 
[Total 10]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
6
(i) Let S be the state space. We say that {π j | j∈S} is a stationary probability
distribution for a Markov chain with transition matrix P if the following
hold for all j∈S :
j i ij
i S
p
∈
π =Σπ , OR π = π P
j 1
j∈S
Σπ = .
πi ≥ 0
(ii) With state space {Working, Broken}
Transition matrix A =
0.95 0.05
0.6 0.4
⎛ ⎞
⎜ ⎟
⎝ ⎠
(iii) This requires the stationary distribution π which satisfies
πA = π
0.95πW + 0.6πB = πW
0.05πW + 0.4πB = πB
and πW + πB =1
πW =12πB
πW =12 /13
πB =1/13
Subject CT4 (Models) – September 2012 – Examiners’ Report
Page 10
So 1 in 13 games is cancelled.
OR
The average number of games for which the lights work before breaking
down is 1/0.05 = 20 games.
Once they have broken down the expected number of games for which
they will be out of action is 1/0.6 = 5/3 games.
Therefore the proportion of games for which the lights are out of action
is
5 / 3 5 1
20 (5 / 3) 65 13
= =
+
So 1 in 13 games is cancelled.
(iv) First we need to find the new stationary probabilities.
Transition matrix A′ =
0.95 0.05
0.8 0.2
⎛ ⎞
⎜ ⎟
⎝ ⎠
0.95πW′ + 0.8πB′ = πW′
and πW′ + πB′ =1
Giving πW′ =16 /17 πB′ =1/17
Lost income (including fees to repair company):
with Floodwatch: ($10,000 + $1,000)*1/13
with Light Fantastic: ($10,000 + X)*1/17 where X is fee to be negotiated.
So need: 11000/13 = (10000 + X)/17
X = $4,384.62 per day.

% In part (i) no credit was given for wordy description of “what happens in the long run”.
% In part (iii) the question said “derive”, so we needed an explanation of where the answer came from: only limited credit was given for writing down a numerical answer (even if correct) without explanation. Moreover, in part (iii) calculating the stationary distribution was not sufficient for full credit: we were looking for the correct element to be identified and its value indicated i.e. an explicit statement that “1 in 13 games is cancelled”. Parts (i), (ii) and (iii) of this question were well answered, and many candidates also evaluated the proportion of games that would be cancelled under the new floodlight regime in part (iv).
% Few were able to compute the daily saving, however.
% Subject CT4 (Models) – September 2012 – Examiners’ Report
% Page 11
\end{document}
