\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}

\usepackage{amsmath}

\usepackage{chngpage}



%\usepackage{bigints}

\usepackage{vmargin}



% left top textwidth textheight headheight



% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}
\setcounter{MaxMatrixCols}{10}
\begin{document}
\begin{enumerate}
PLEASE TURN OVER7
Mr Bunn the baker made 12 pies to sell in his shop. He placed the pies in the shop at
9 a.m. During the rest of the day the following events took place.
Time
10 a.m.
11 a.m.
12 noon
1 p.m.
2 p.m.
3 p.m.
5 p.m.
8
Event
A boy bought two pies
A man bought three pies
Mr Bunn accidentally sat on one pie and squashed it so it could not be
sold
A woman bought two pies
A dog from across the street ran into Mr Bunn’s shop and stole two
pies
A girl on the way home from school bought one pie
Mr Bunn closed for the day and the remaining pie was still in the shop
(i) Estimate the time it takes Mr Bunn to sell 40% of the pies he makes, using the
Nelson-Aalen estimator.
[6]
(ii) Comment on whether you think this estimate would be a good basis for Mr
Bunn to plan his future production of pies.



%%%%%%%%%%%%%%%%%%%%%

7
(i)
The sequence of events described may be summarised in the table below
Duration t j Pies in shop
n j Pies bought
d j Pies destroyed or
stolen, c j
1
2
3
4
5
6 12
10
7
6
4
2 2
3
0
2
0
1 0
0
1
0
2
0
The hazard of pies being bought is thus
2/12 at duration
3/10 at duration
2/6 at duration
1/2 at duration
1 hour
2 hours
4 hours
6 hours
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Nelson-Aalen estimate of the survival function, $S(t)$, is then
Duration Nelson-Aalen estimate of S(t)
0 \leq t < 1
1\leq t < 2
2 \leq t < 4
4 \leq t < 6
6 \leq t < 8 1
exp [-2/12] = 0.8465
exp [-(2/12 + 3/10)] = 0.6271
exp [-(2/12 + 3/10 + 2/6)] = 0.4493
exp [-(2/12 + 3/10 + 2/6 + 1/2)] = 0.2725
%%%%%%%%%----------------------------------%
The Nelson-Aalen estimate is a step function.
We need t for which $S(t) = 0.6$.
Therefore it will be 4 hours until Mr Bunn has sold 40% of his pies.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
(ii)
\begin{itemize}
\item The estimate would not be a good basis on which to plan future production.
And how long it takes to sell 40\% of your goods is not very relevant for future
production.
\item It is based on only one day’s experience, and a good basis for future production should be based on several days, probably iinvolving different days of the week.
Sales of pies may vary seasonally: data from a winter’s day may tell Mr Bunn little about the demand for pies in summer.
\item Mr Bunn might be more careful in future not to sit on his pies, and might take steps to avoid the dog from across the street stealing pies.
\item The proportion of pies sold will depend on the number of pies Mr Bunn stocks. He should not assume if he had twice as many pies he would still sell 40\% of them in 4 hours.
\item Mr Bunn may vary his sales strategy, by, for example, reducing his prices The method does, however, take account to of censored data.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
In part (i) the question said “estimate”, so some indication of how the answer was arrived at
was necessary, although not every detail was required. As a bare minimum full credit could
be obtained for first three hazards at times 1, 2 and 4, some statement of what the Nelson-
Aalen estimate of S(t) is, the fact that we are looking for S(t) = 0.6, and some numbers to
demonstrate that S(t) = 0.6 happens at duration 4 hours. Answers that used the logarithm of
S(t) were acceptable. Answers to part (ii) were encouraging. A substantial proportion of
candidates made sensible points.


\end{document}
