\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}
% left top textwidth textheight headheight

% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}
\setcounter{MaxMatrixCols}{10}

\begin{document}



\begin{enumerate}

1 Give two examples of the main types of liability insurance, stating for each example a typical insured peril.

2 A claim amount distribution is normal with unknown mean \mu and known standard deviation \$50. Based on past experience a suitable prior distribution for \mu is normal with mean \$300 and standard deviation \$20.
(i)
(ii)
Calculate the prior probability that \mu, the mean of the claim amount distribution, is less than \$270.

A random sample of 10 current claims has a mean of \$270.
(a) Determine the posterior distribution of \mu.
(b) Calculate the posterior probability that \mu is less than \$270 and
comment on your answer.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
1
Type
Employers’ liability
Motor 3 rd party liability
Public liability
Product liability
Professional Indemnity
Typical perils
\item  Accidents caused by employer negligence
\item  Exposure to harmful substances
\item  Exposure to harmful working conditions
\item  Road traffic accidents
\item  Will relate to the type of policy
\item  Faulty design, manufacture or packaging of
product
\item  Incorrect or misleading instructions
\item  Wrong medical diagnosis, error in medical
operation etc.
Comment: Most of the candidates did very well here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
2
(i)
P ( \mu < 270) = P ( N (300, 20 2 ) < 270)
270 − 300
)
20
= P ( N (0,1) < − 1.5)
= P ( N (0,1) <
= 1 − 0.93319
= 0.06681
(ii)
(a)
Using the result from page 28 of the tables, the posterior distribution of \mu is
normal, with mean
10 × 270 300
+ 2 )
2
20
\mu * = 50
(
10
1
( 2 + 2 )
50
20
= \$281.54
and variance
\sigma * 2 =
1
10
1
+ 2
2
50
20
= 153.85 = 12.40 2
So the posterior distribution of \mu is N (281.54,12.40 2 ) .
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(b)
The posterior probability required is given by:
270 − 281.54
)
12.4
= P ( N (0,1) < − 0.931)
= 1 − (0.9 × 0.82381 + 0.1 × 0.82639)
P ( N (281.54,12.40 2 ) < 270) = P ( N (0,1) <
= 0.1759
Comment: The probability that the true mean is less than \$270 has risen, as the sample evidence suggests that the true mean is less than the mean of the prior distribution. Nevertheless, the sample size is relatively small, and the variance of the prior distribution is also small, so that a reasonable weight is still given to the prior information.
Comment: Some candidates did not use tables for generating the posterior
parameters in (ii)(a). A few gave the correct interpretation in (ii)(b).

\end{document}
