\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}
% left top textwidth textheight headheight

% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}
\setcounter{MaxMatrixCols}{10}

\begin{document}An insurer is considering whether to outsource its advertising. If it decides not to
outsource, it expects to spend £1,400,000 next year on its advertising, and believes
that this would result in a portfolio of 100,000 policies. It has received quotations
from two different companies for outsourcing its advertising.
Company A would cost £2,100,000 per year, and believes that this would result in the
business expanding to 125,000 policies.
Company B would cost £3,000,000 per year, and believes that this would expand the
business to 140,000 policies.
At present, each policy returns a profit to the company of £30 per year; but this is not
guaranteed in the future. The company has assessed that it will stay at this level with
probability 0.6, but could reduce to £20 per year with probability 0.25, or increase to
£40 per year with probability 0.15.
(i) Explain which of the three options can be immediately discarded.
(ii) Determine the Bayes solution to the problem of maximising the profit to the
company over the coming year.

3
(i) Company B can be discounted immediately because it is dominated by both of
the other two options.
(ii) Profit Table (in millions)
Current
Company A
£20
0.6
0.4
£30
£40
1.6
2.6
1.65 2.9
E[Profit]
1.5
1.525
The insurer should choose Company A.
