\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chng%%-- Page}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
2
Two poachers can escape from either the farmyard or the fields of an estate. Two gamekeepers are employed by the estate to catch them. The gamekeepers have three options: both patrol the farmyard (action a 1 ); one of them patrols the farmyard and the
other patrols the fields (action a 2 ); both gamekeepers patrol the fields (action a 3 ). The poachers also have three options: both try to escape through the farmyard (option \theta 1 ); one of them tries to escape from the farmyard and the other from the fields (option \theta 2 ); both try to escape over the fields (option \theta 3 ). The number of birds they manage to poach under each combination of \theta i and a j can be found in the following table:

h combination of θ i and a j can be found in the following table:
θ 1
θ 2
θ 3
a 1 a 2 a 3
0
75
120 90
0
90 120
75
0

Assume the goal of the gamekeepers and the poachers is to minimise and maximise
the number of birds poached, respectively.
\begin{enumerate}
\item (i) Show that \theta 1 does not dominate \theta 2 and vice versa.
\item (ii) Determine the minimax solution to this problem for the gamekeepers, and
state the maximum number of birds that will be poached if they adopt this
strategy.
\item Given the prior distribution P(\theta 1 ) = 0.25, P(\theta 2 ) = 0.35 and P(\theta 2 ) = 0.4,
determine the Bayes solution to the problem for the gamekeepers.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
2
(i)
\theta  1 would dominate \theta  2 (and vice versa) if the amount of birds poached under
each of a 1 , a 2 and a 3 is higher in each case for \theta  1 (\theta  2 )
\theta  1 provides a better outcome for a 2 (90 > 0) and a 3 (120 > 75)
\theta  2 provides a better outcome for a 1 (0 < 75)
(ii)
Minimax solution for gamekeepers. Worst loss under each option would be:
\begin{itemize}
\item a 1 = max (0,75,120) = 120
\item a 2 = max (90,0,90) = 90
\item a 3 = max (120,75,0) = 120
\end{itemize}

Therefore, gamekeepers would choose a 2 and lose 90 birds.
(iii)
\begin{itemize}
\item a 1 = 0.25 \times 0 + 0.35 \times 75 + 0.4 \times 120 = 74.25
\item a 2 = 0.25 \times 90 + 0.35 \times 0 + 0.4 \times 90 = 58.5
\item a 3 = 0.25 \times 120 + 0.35 \times 75 + 0.4 \times 0 = 56.25
\end{itemize}
Hence the Bayes decision is a 3 .
\end{document}
