\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}

%%--- Question 9
\item An electronics company developed a revolutionary new battery which it believed
would make it enormous profits. It commissioned a sub-contractor to estimate the
survival function of battery life for the first 12 prototypes. The sub-contractor
inserted each prototype battery into an identical electrical device at the same time and
measured the duration elapsing between the time each device was switched on and the time its battery ran out. The sub-contractor was instructed to terminate the test
immediately after the failure of the 8th battery, and to return all 12 batteries to the
company.
When the test was complete, the sub-contractor reported that he had terminated the
test after 150 days. He further reported that:
•
•
•
•
two batteries had failed after 97 days
three further batteries had failed after 120 days
two further batteries had failed after 141 days
one further battery had failed after 150 days
However, he reported that he was only able to return 11 batteries, as one had exploded
after 110 days, and he had treated this battery as censored at that duration when
working out the Kaplan-Meier estimate of the survival function.
(i) State, with reasons, the forms of censoring present in this study.

\item (ii) Calculate the Kaplan-Meier estimate of the survival function based on the
information supplied by the sub-contractor.

In his report, the sub-contractor claimed that the Kaplan-Meier estimate of the
survival function at the duration when the investigation was terminated was 0.2727.
\item (iii)
Explain why the sub-contractor’s Kaplan-Meier estimate would be consistent
with him having stolen the battery he claimed had exploded.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
[Total 11]
CT4 S2009—610
An investigation into the mortality of men engaged in a hazardous occupation was
carried out. The following is an extract from the results.
Age x Initial
exposed-to-risk E x
30
31
32
33
34
35
36
37
38 950
1,200
1,200
900
1,000
1,100
800
1,250
1,400
Observed
deaths θ x
12
14
16
9
11
15
10
16
17
q ˆ x
0.0126
0.0117
0.0133
0.0100
0.0110
0.0136
0.0125
0.0128
0.0121
It was decided to graduate the results with reference to English Life Table 15 (males).
o
The formula used for the graduation was q x = 10 q x s .
(i) Using a test of the overall fit of the graduated rates to the data, test the
hypothesis that the underlying mortality of men in the hazardous occupation is
in accordance with the graduation formula given above.
[6]
\item (ii) Test the graduation using two other tests which detect different features of the
graduation. For each test you apply:
(a)
(b)
(c)
State the feature of the graduation it is designed to detect.
Carry out the test.
State your conclusion.
[7]
[Total 13]
CT4 S2009—7
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
9
(i) Type II censoring as the study was terminated after a pre-determined number
of failures. Random censoring of the device which exploded.
\item (ii) According to the information supplied by the sub-contractor, the Kaplan-
Meier estimate of the survival function should be calculated as follows:
Page 11
j t j N j
0 0 12
1 97 12
d j c j d j /N j 1 – d j /N j
2 1 2/12 10/12  — %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% — Examiners’ Report
2 120 9 3 0 3/9 6/9
3 141 6 2 0 2/6 4/6
4 150 4 1 3 1/4 3/4
The Kaplan-Meier estimate is then
 d j
S ˆ ( t )    1 
 N j
t j  t 




so we have
t S ˆ ( t )
0  t  97 1
97  t  120 5/6
120  t  141 5/9
141  t  150 10/27
150  t 5/18 = 0.2778
\item (iii)Since 5/18 is not equal to 0.2727, the sub-contractor‟s story is internally
inconsistent. The Kaplan-Meier estimate of the survival function after the
failure of the 8th battery of 0.2727 would be obtained had only 11 batteries
been tested at the start, and no battery being censored, as shown in the
following table.
d j c j d j /N j 1 – d j /N j
11 2 0 2/11 9/11
120 9 3 0 3/9 6/9
3 141 6 2 0 2/6 4/6
4 150 4 1 0 1/4 3/4
+1⁄2 +1⁄2
j t j N j
0 0 11
1 97 2
The Kaplan-Meier estimate is then
 d j
S ˆ ( t )    1 
 N j
t j  t 




so we have
t
Page 12
S ˆ ( t )  — %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% — Examiners’ Report
0  t  97 1
97  t  120 9/11
120  t  141 6/11
141  t  150 4/11
150  t 3/11 = 0.2727
Therefore the value of S ˆ (150) reported by the sub-contractor is consistent with
him having stolen the last battery.
Many candidates scored highly on this question. Credit was given in part (i) for other
types of censoring provided that a sensible reason was given. In part \item (iii), for full
credit some kind of calculation of an alternative survival function was needed,
together with an explanation of why this provided evidence to support the suggestion
that the sub-contractor has stolen the battery.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

10
(i) The chi-squared test is for the overall fit of the graduated rates to the data
The test statistic is
 z x 2 , where
o
z x 
 x  E x q x
o
o
.
E x q x (1  q x )
o
The calculations are shown in the table below (since q x is
o
small we use the approximation z x 
 x  E x q x
o
.
E x q x
Page 13
 x o o
Age x q x E x q x
30 12 0.0091 31 14 0.0094 32 16 33
8.645
z x
z x 2
1.141 1.302
11.28 0.810 0.656
0.0097 11.64 1.278 1.633
9 0.0099 8.91 0.030 0.001
34 11 0.0106 10.60 0.123 0.015
35 15 0.0116 12.76 0.627 0.393
36 10 0.0127 10.16 -0.050 0.003
37 16 0.0138 17.25 -0.301 0.091  — %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% — Examiners’ Report
38
17
0.0149
20.86
-0.845
0.714
∑
4.808
The test statistic has a chi-squared distribution with degrees of freedom (d.f.)
given by number of ages
o
– 1 (for parameter of function linking q x and q x s )
– some d.f. for constraints imposed by choice of standard table
The critical value of the chi-squared distribution is
11.07 with 5 d.f.
12.59 with 6 d.f.
14.07 with 7 d.f.
15.51 with 8 d.f.
16.92 with 9 d.f. at the 5\% level (from tables)
Since 4.808 < 11.07 (or 12.59 etc.) there is no evidence to reject the null
hypothesis that the graduated rates are the true rates underlying the crude
rates.
\item (ii)
EITHER
Signs test
a. The Signs test looks for overall bias.
b. The number of positive signs among the z x s
is distributed Binomial (9, 0.5).
We observe 6 positive signs.
The probability of obtaining 6 or more positive signs is
(from tables)
1 – 0.7461 = 0.2539.
[Alternatively, candidates could calculate the probability of obtaining exactly 6
positive signs, which is 0.1641]
Since this is greater than 0.025 (two-tailed test)
c. we cannot reject the null hypothesis and we conclude that the
graduated rates are not systematically higher or lower than the crude
rates.
OR
Cumulative Deviations test
Page 14  — %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% — Examiners’ Report
a. When applied over the whole age range, the Cumulative Deviations
test looks for overall bias
b. The test statistic is

o

    x  E x q x  
x
o
 Normal(0,1)
 E x q x
x
o
o
Age x  x 30 12 31 14 11.28 2.72
32 16 11.64 4.36
33 9 8.91 0.09
34 11 10.60 0.40
35 15 12.76 2.24
36 10 10.16 -0.16
37 16 17.25 -1.25
38 17 20.86 -3.86
∑ 112.105 7.895
E x q x
8.645
So the value of the test statistic is
 x  E x q x
3.355
7.895
 0.7457
112.105
Using a 5\% level of significance, we see that
1.96 < 0.7457 < 1.96
c. We accept the null hypothesis at the 5\% level of significance and
conclude there is no overall bias in the graduation.
Grouping of Signs test
a. The Grouping of Signs test looks for runs or clumps of deviations of
the same sign OR the grouping of signs test tests for overgraduation.
b. We have:
9 ages in total
6 positive deviations
3 negative deviations
We have 1 positive run
Pr[1 positive run] is therefore equal to
Page 15  — %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% — Examiners’ Report
 5  4 
  
4
4
 0   1  

 0.0476
 9 
 9.8.7  84


 
 3.2 
 6 
Since this is less than 0.05 (using a one-tailed
test)
c. We reject the null hypothesis that the graduated rates are the true rates
underlying the crude rates (OR we conclude that the graduation is
unsatisfactory OR there is evidence of over-graduation).
Individual Standardised Deviations test
a. The Individual Standardised Deviations tests looks for individual large
deviations at particular ages.
b. If the graduated rates were the true rates underlying the observed rates
we would expect the individual deviations to be distributed Normal (0,1) and therefore only 1 in 20 z x s should have absolute magnitudes
greater than 1.96. Looking at the z x s we see that the largest individual
deviation is 1.278. Since this is less in absolute magnitude than 1.96
c. we cannot reject the null hypothesis that the graduated rates are the
true rates underlying the crude rates.
Answers to this question were disappointing compared with previous years. A
common error was for candidates to misread the question and to try to compare the
observed number of deaths with an ‘expected’ number computed on the basis of the
^
q x given in the question. These candidates were, in effect, examining deviations
based solely on rounding! Candidates who made this error were penalised in part (i),
but could gain credit for some of the alternative tests in part \item (ii) provided that they
performed the tests correctly.
\end{document}
