\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chng%%-- Page}
%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}
\setcounter{MaxMatrixCols}{10}
\begin{document}
1 An actuary is considering a portfolio of insurance policies. This portfolio only contains five policies. Each policy has been taken out by a different couple who are getting married on the same day in the same city. Each policy would pay out in the
event of rain on that day in that city.
(i) Explain why these policies may not meet the ideal criteria for an insurable risk. 
(ii) Suggest two ways in which either the portfolio or the policy terms could be changed in order to better meet the criteria for insurable risks. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
Solutions
Q1 (i) Not sufficiently diversified / not independent / not large enough sample 
Not easily quantifiable 
Risk is potentially quite large (not necessarily a remote risk) 
Policyholders may not have financial interest in risk 
Moral hazard 
[Max 3]
(ii) Increase number of policies 
ensure definition of bad weather is extreme 
Diversify policies between different cities 
Diversify policies between different days 
Introduce policy excess 
[Max 2]

Most candidates were able to relate the theory of insurable risks to this question. A few candidates failed to sufficiently distinguish between their points, or did not consider the actual context presented.


\end{document}
