\documentclass[]{report}

\voffset=-1.5cm
\oddsidemargin=0.0cm
\textwidth = 480pt

\usepackage{framed}
\usepackage{subfiles}
\usepackage{graphics}
\usepackage{newlfont}
\usepackage{eurosym}
\usepackage{amsmath,amsthm,amsfonts}
\usepackage{amsmath}
\usepackage{color}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\begin{document}
	
\section{Question D2 - Binomial Distribution (2 Marks) } % 10 Marks
A biased coin yields `Tails' on $48\%$ of throws. Consider an experiment that consists of throwing this coin 11 times.
\begin{itemize}
	\item[a.] (1 Mark) Evaluate the following term $^{11}C_2$.
	\item[b.] (1 Mark) Compute the probability of getting two `Tails' in this experiment.
\end{itemize}

\section{Question D1 - Binomial Distribution } % 10 Marks
An inspector of computer parts selects a random sample of components
from a large batch to decide whether or not to audit the full batch.

%---------------------------------------------%
\begin{itemize}
	\item[(i)] lf 20\% or more of the sample is defective, the entire batch is
	inspected, Calculate the probability of this happening if it is
	thought that the population contains 4\% defective components and
	a sample of 20 is selected.
	\item[(ii)] lf 10\% or more of the sample is defective, the entire batch is
	inspected. Calculate the probability of this happening if it is
	thought that the population contains 4\% defective components and
	a sample of 50 is selected.
	(10 marks)
\end{itemize}


\section{Question D2 - Binomial Distribution (2 Marks) } % 10 Marks
Under what circumstances is it appropriate to use the binomial distribution when calculating probabilities?					(1 mark)

(b) 	Flextronics supply PCB boards to Dell.  You are a production manager with Dell.  There is a constant probability of 0.01 that a board will be defective.  You select 20 boards at random.  What is the probability that:
\begin{itemize}
	\item[(i)]	0 boards will be defective
	\item[(ii)]	1 or more boards will be defective
	\item[(iii)]	2 or less boards will be defective			
	(6 marks)
\end{itemize}

\end{document}	