Question 1
A fair coin is tossed three times. Write out the sample space S.

Write the sample points of the following events

(A) the event that the first toss is heads
(B) the event that the second toss is heads
(C) exactly two heads in a row.

Question 2
A couple has two children; the sample space is S = {bb, bg, gb, gg} with probability 0.25 for each sample point. Find the probability p that both children are boys if it is known that: 
(i) at least one of the children is a boy;
(ii) the older child is a girl. 		(Revision for 2012: older child is a boy)
Question 3
Compute the probability p of each event:
(a) An even number appears in the toss of a fair die;
(b) One or more tails appear in the toss of three fair coins;
(c) A blue marble appears in a random drawing of one marble from a box containing four white, three blue, and five red marbles.

Question 4
Suppose a student is selected at random from 100 students where 30 are taking mathematics, 20 are taking computer science, and 10 are taking mathematics and computer science. Find the probability p that the student is taking mathematics or computer science.
Question 5
A lot contains 13 items of which 4 are defective. Three items are drawn at random from the lot one after the other. Find the probability p that all three are nondefective.
Question 6
The probability that A hits a target is 1/3 and the probability that B hits a target is 1/5.  They both fire at the target. Find the probability that:

(a) A does not hit the target; 
(b) both hit the target; 
(c) one of them hits the target; 
(d) neither hits the target.

Question 7
On completion of a programming project, four programmers from a team submit a collection of subroutines to an acceptance group. The following table shows the percentage of subroutines each programmer submitted and the probability that a subroutine submitted by each programmer will pass the certification test based on historical data.

Programmer	A	B	C	D
Proportion of subroutines submitted	0.10	0.20	0.30	0.40
Probability of acceptance	0.55	0.60	0.95	0.75

(i) What is the proportion of subroutines that pass the acceptance test?
(ii) After the acceptance tests are completed, one of the subroutines is selected at random and found to have passed the test. What is the probability that it was written by Programmer A?

Question 8
Two manufacturing plants produce similar parts. Plant A produces 1000 parts, 100 of which are defective. Plant B produces 2000 parts, 170 of which are defective. 

1.	Present this information using a contingency table

2.	A part is selected at random and found to be defective. What is the probability that it came from plant A ?

Question 9.
A machine is composed of 2 components A and B, which function independently of one another. The machine works only if both components are working. The probability that component A works is 0.98 and the probability that component B works is 0.95.
•	What is the probability that the machine works, 
•	What is the probability that both components have failed.
 

Question 10.  
The random variables X and Y have the following joint distribution
 
	Y=0	Y=1 
	Y=2
X=0	2/10	1/10	2/10
X=1	0/10	2/10	1/10
X=2	1/10	0/10	1/10

 
i)	Calculate the marginal distributions of X and Y.     
ii)	What is the expected value of X?
iii)	What is the expected value of Y?                 
 


%====================================================================================================%


Question 1 

Combinations and Permutations Compute the following:
 

 
1.	5C2
2.	5P2
3.	4C0
4.	4C4
5.	6C1
6.	6P3
7.	6C2
8.	6C3
 


 
Question 2

Six horses are entered in a race. Larry picks two of the horses at random, and bets on them. 
How many ways are there of picking two horses at random from six?

Question 3

Suppose a fair coin is tossed six times. The number of heads which can occur with their respective
probabilities follows:

xi	0	1	2	3	4	5	6
p(xi)	1/64	6/64	15/64	20/64	15/64	6/64	1/64

1.	Compute the expected value (i.e. expected number of heads).
2.	Compute the variance of the heads.
Question 4

A player tosses two fair coins. He wins $2 if two heads occur, and $1 if one head occurs. On the other hand, he loses $3 if no heads occur. Find the expected value E(X) of the game. Is the game fair? 

(The game is fair, favorable, or unfavorable to the player according as E(X) = 0,E(X) > 0 or E(X) < 0.)

Question 5

A lottery with 500 tickets gives one prize of $100, three prizes of $50 each, and five prizes of $25 each.
(a) Find the expected winnings of a ticket. 
(b) If a ticket costs $1, what is the expected value of the game?

Question 6
A random sample with replacement of size n = 2 is drawn from the set {1, 2, 3}, yielding the following
9-element sample space:

S = {(1, 1), (1, 2), (1, 3), (2, 1), (2, 2), (2, 3), (3, 1), (3, 2), (3, 3)}

Each pairing is equally probable. Let X denote the sum of the two numbers. 

Construct a table that contains all the possible values of X and the probability of that outcome.

Compute the expected value E(X).

Question 7

Suppose X is a random variable with mean 15 and standard deviation 5.
Compute the expected value of X2  (Recall: standard deviation is the square root of variance)



%====================================================================================================%



Question1 : Discrete Random Variables
Consider a binomial experiment where the number of independent trials is 10, and the probability of success is 0.25. Let X denote the number of successes in the 10 trials.
1)	Write out the sample space of X.
2)	For the each of the following events, write out the sample points:
a.	the number of successes is less than or equal to 3,
b.	the number of successes is greater than 4, 
c.	the number of successes is greater than 8, 
d.	the number of successes is between 4 and 7 inclusive.
Question 2: Discrete Random Variables
By considering the relevant sample points from the sample space in question 1, show that 
1)	P(4 ≤X ≤8) = P(X ≥4) – P(X ≥9 )
2)	P(X=4) = P(X ≥4) – P(X ≥5 )
3)	P(X≤7) = 1- P(X ≥8)
Question 3:  Poisson Distribution
Cars enter a car wash at an average rate of 4 per half hour. 
1)	Compute the probability that the 3 cars arrive in a half hour period.
2)	Compute the probability that the 6 cars arrive in a one hour period.
Question 4 : Binomial Distribution
Let X denote the number of bits received in error in a digital communication channel, and assume that X is a binomial random variable with p=0.10. If 100 bits are transmitted, use the binomial tables to determine the following
1)	P(X=10)
2)	P(X ≥10)
3)	P(X ≤20)
Question 5 : Binomial Distribution
In a test of a printed circuits board using a random test pattern,  an array of 10 bits is equally likely to be “zero” or “one”. Assume that the bits are independent. 
What is the probability that all bits are “ones”?  (Answer : 0.0010)
Question 6 : Binomial Distribution
An examination contains 20 multiple choice questions with four choices per questions. A pass is obtained by answering  10 questions correctly. Calculate the probability that a student who chooses the answer to each question at random will pass the examination.  ( Answer : 0.0139 )

Question 7 : Binomial Distribution
An electronics product contains 20 integrated circuits. The probability that any integrated circuit is defective is 0.20. The ICs operate independently of each other.  The product operates only if there are no defective ICs. What is the probability that the product operates?
It is not economically viable to repair the product if there are more than 3 ICs are defective.  What is the probability that it will not be repaired if broken.
Question 8 : Poisson Distribution
Phone calls arrive at a help desk at the rate of 48 per hour.
1)	Find the probability of receiving three calls in a five minute period of time.
2)	What is the probability of receiving 5 or fewer calls in a 15 minute interval of time.
Question 9: Poisson Distribution
The number of particles emitted by a radioactive source in 1 minute has a Poisson distribution with parameter m=1.2, called the emission rate. Calculate the probability that
1)	No particle is emitted
2)	At least 4 particles are emitted
Question 10 : Poisson Approximation of Binomial distribution
A particular electronic component is produced by a manufacturing process for which 0.1% of components are produced are defective. Consider a consignment of 5,000 of these times.
Let X denote the number of defective components in the consignment.
Use the Poisson Approximation of the binomial distribution to calculate P(X ≤5)









%====================================================================================================%

MA4413 Tutorial for Week 6 : Continuous Probability Distributions
Question 1:  Use the Murdoch Barnes Tables to compute each of the following. 
(Recall the complement, symmetry and interval rules).

 
I.	 
II.	 
III.	 
IV.	 
V.	 
VI.	 
VII.	 
VIII.	 
 

Question2:  The lifetime of an electrical component is known to follow a normal distribution with a mean of 2,000 hours and a standard deviation of 200 hours.  

Compute the probability that a randomly selected component will last 
1.	more than 2,220 hours, 
2.	between 2,000 and 2,400 hours. 

Question 3
The amount of time required for routine automobile transmission service is normally distributed with the mean 45 minutes and the standard deviation 8.0 minutes. 

The service manager plans to have work begin on the transmission of a customer’s car 10 min after the car is dropped off, and the customer is told that the car will be ready within 1 hour total time (i.e. after  the car is dropped off). 

What is the probability that the service manager will be wrong?

Question 4
 The mass of shire horses is assumed to have a normal distribution with mean 1000kg and standard deviation 50kg. 
I.	Calculate the probability that the mass of a shire horse is more than 975kgs.
II.	Calculate the probability that the mass of a shire horse is between 945kg and 1032kg.                    
III.	What weight is exceeded by 2.5% of shire horses?                 

Question 5
The average lifespan of a PC monitor is 6 years. You may assume that the lifespan
of monitors follows an exponential probability distribution.

I.	What is the probability that the lifespan of the monitor will be at least 5 years?
II.	What is the probability that the lifespan of the monitor will not exceed 4 years?
III.	What is the probability of the lifespan being between 5 years and 7 years?

[ For  exponential distributions, with mean duration μ:   P(X ≤ k) = 1- e(-k/μ) ]







