\documentclass[]{report}

\voffset=-1.5cm
\oddsidemargin=0.0cm
\textwidth = 480pt

\usepackage{framed}
\usepackage{subfiles}
\usepackage{graphics}
\usepackage{newlfont}
\usepackage{eurosym}
\usepackage{amsmath,amsthm,amsfonts}
\usepackage{amsmath}
\usepackage{color}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\begin{document}
	
\section{Question D3 -  Poisson Distribution (2 Marks) }  % 12 Marks
Suppose that a telephone help-line receives 4 calls per hour during offices hours.
\begin{itemize}
	\item[a.] (1 Mark) Compute the value of $m$ for a 30 minute period during office hours.
	\item[b.] (1 Mark) Compute the probability of the help-line getting exactly one call in a 30 minute period during office hours.
\end{itemize}




\section{Question D05 - Exponential Distribution } % 10 Marks
(c)	Flaws occur in an LCD display at the rate of 0.5 per square mm.  Calculate the probability that:
\begin{itemize}
	\item[(i)]	exactly 2 flaws will occur in a square mm section
	\item[(ii)]	exactly 3 flaws will occur in a 5 square mm section
	\item[(iii)]	5 or more flaws will occur in a 10 square mm section
	(9 marks)
\end{itemize}

(d)	There is a constant probability of 0.05 that the power supply in a server network will fail.  You are required to calculate the probability that the power supply will fail the 4th time it is switched on.
(4 marks)



\section{Question D01  - Poisson Example}
A motor dealership which specializes in agricultural machinery sells on vehicle every 2 days, on average

In this question the unit period is one day. The company expects to sell, on average, 0.5 vehicles every day.
The Poisson mean $m$ is therefore 0.5.

$P(X \geq 1)$

Go to your Poisson tables, and search for the $m=0.5$ column.
We are interested in the probability of \textbf{exactly} one vehicle sold on a particular day.
From the tables we can easily work out P($X \geq 1$), but this is probability of one or more vehicles being sold.
This is not the same thing.
\[P(X \geq 1) = P(X =1) + P( X=2) + P(X=3) + \ldots
P(X \geq 1) = P(X=1) + P(X \geq 2)\]
From tables
$P(X \geq 1)$
$P(X \geq 2)$

Six day working week?
our unit period is now six days.
How many vehicles do we expect to sell in 6 days?
answer = 3
$m=3$

P($X\geq 4$)

\begin{eqnarray}
e^{-6/5} = 0.3011942
e^{-4/5} = 0.449329
e^{-5/5} = 0.3678794
\end{eqnarray}









%---------------------------------------------------%

(c)	XYZ Ltd supplies motherboards to Dell.  You are a production manager with Dell.  There is a constant probability of 0.4 that a board will be defective.  You select 100 boards at random.  Using the log tables for the binomial distribution what is the probability that

\begin{itemize}
	\item[(i)] 0 boards will be defective?
	\item[(ii)] 2 or more boards will be defective?
	\item[(iii)]	5 or less boards will be defective?
\end{itemize}
% (6 marks)
(d)	Use the normal approximation to the binomial to answer (i), (ii) and (iii) in part (c) above.
%(6 marks


\end{document}