\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
PLEASE TURN OVER11
A new disease has been discovered which is transmitted by an airborne virus.
Anyone who contracts the disease suffers a high fever and then in 60% of cases dies
within an hour and in 40% of cases recovers. Having suffered from the disease once,
a person builds up antibodies to the disease and thereafter is immune.
\begin{enumerate}[(a)]
\item (i) Draw a multiple state diagram illustrating the process, labelling the states and
possible transitions between states.
\item 
(ii) Express the likelihood of the process in terms of the transition intensities and
other observable quantities, defining all the terms you use
\item 
(iii) Derive the maximum likelihood estimator of the rate of first time sickness.
\smallskip
Three years ago medical students visited the island where the disease was first
discovered and found that of the population of 2,500 people, 860 had suffered from
the disease but recovered. They asked the leaders of the island to keep records of the
occurrence and the outcome of each incidence of the disease. The students intended
to return exactly three years later to collect the information.
\item 
(iv) Derive an expression (in terms of the transition intensities) for the probability
that an islander who has never suffered from the disease will still be alive in
three years’ time.
\item 
(v) Set out the information which the students would need when they returned three years later in order to calculate the rate of sickness from the disease. 
\end{enumerate}
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
11
ALTERNATIVE 1
\begin{itemize}
\item (i)
0.4ρ
Healthy
(H)
Recovered and
immune
(R)
0.6ρ
Dead from other
reason (O)
Dead from disease
(D)
(ii)
The likelihood is
 
L  exp  0.6   0.4   
 0.4  
d HR
     
HO
d HO
RO
HO
   exp        0.6  
H
RO
R
d RO
where
the four states are H – healthy, R – recovered, D – dead from
the disease and O – dead for some other reason, and
 I is the waiting time in state I (I = H, R)
Page 16
d HD %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% – April 2015 – Examiners’ Report
d IJ is the number of transitions from state I to state J (J = R, D, O)
and  IJ is the intensity of the transition from state I to state J and ρ is
the rate of first time sickness
\item (iii)
Taking logarithms of the likelihood we have:
ln L    0.6   0.4    H  d HS ln  0.6    d HR ln  0.4   plus terms not
dependent on 
Differentiating with respect to  gives:
d  ln L 
d 
0.6 d HS 0.4 d HR
  

0.6 
0.4 
H
and setting this to zero gives a maximum likelihood estimate of 
 ˆ 
d HD  d HR
 H
This is a maximum as the second derivative
d 2  ln L 
 d   2

d HD  d HR
   2
must
be negative.
\item (iv)
We want a person in H at t = 0 not dead at t = 3.
So they can either be healthy throughout:
HH/
t p 0
  3
 
  3
 
HO
 exp    0.4   0.6    du   exp       HO du 
  0
 
  0
 




or go H to R and stay there which is the integral between
0 and 3 of the product of
 
 
HO
u
surviving healthy for a period u i.e. exp  
getting sick and recovering at time u i.e. 0.4ρdu
  

RO
 3  u     ,
and staying recovered i.e. exp    

where we ignore the short time spent in the sick state.
3
 
 
  

\item So altogether exp      HO u  0.4  exp    RO  3  u   du .
 
 
 
 

0
Page 17 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% – April 2015 – Examiners’ Report
\item (v)
We need enough information to calculate the number of transitions from
Healthy to Sick and the waiting time in the “Healthy” state. Hence we will
need:
Date of birth of all births in last three years.
Date of death of all “Healthy” deaths in last three years.
Date of any other immigration or emigration of Healthy people in last three
years.
Date at which each person who contracted the disease fell ill
\item ALTERNATIVE 2
(i)
Suffering
from
disease
(S)
Healthy
(H)
Recovered and
immune
(R)
Dead
(D)
(ii)
The likelihood is
 
       
d
d
d
d
HD
RD
SR
SD




       
   
L  exp  HS   HD  H exp  RD  R exp  SR   SD  S  HS
HD
RD
SR
SD
where
 I is the waiting time in state I,
d IJ is the number of transitions from state I to state J,
and  IJ is the intensity of the transition from state I to state J.
\item The four states are H – healthy, S – suffering from disease,
R – recovered, D – dead.
Page 18
d HS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% – April 2015 – Examiners’ Report
(iii)
Taking logarithms of the likelihood we have:
 
ln L   HS  H  d HS ln  HS
plus terms not dependant on  HS
\item Differentiating with respect to  HS gives:
d  ln L 
d  HS
  H 
d HS
 HS
and setting this to zero gives a maximum likelihood estimate of  HS
 ˆ
HS

d HS
 H
This is a maximum as the second derivative
d 2  ln L 
 d  
HS
2

d HS
  
HS
2
must be
negative.
\item (iv)
We want a person in H at t = 0 neither suffering from the
disease nor dead at t = 3.
So they can either be healthy throughout:
HH
t p 0
  3
 
HS
HD
 exp       du 
  0
 


or to go H to S to R and stay there which is the integral between
0 and 3 of the product of:
 
 
HS
HD
u
surviving healthy for a period u i.e. exp   
getting sick at time u i.e.  HS du and recovering, i.e. 0.4
  
    ,
RD
 3  u 
and staying recovered i.e. exp    

where we ignore the short time spent in the sick state.
3
 
 
  

So altogether exp    HS   HD u   HS 0.4 exp    RD  3  u   du .
 
 
 
 

0
Page 19 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% – April 2015 – Examiners’ Report
\item (v)
We need enough information to calculate the number of transitions from
Healthy to Sick and the waiting time in the “Healthy” state. Hence we will
need:
Date of birth of all births in last three years.
Date of death of all “Healthy” deaths in last three years.
Date of any other immigration or emigration of Healthy people in last three
years.
Date at which each person who contracted the disease fell ill.
Various alternative (usually simpler) models were suggested by some candidates. For
example a much simpler model was proposed with only three states:
Healthy
0.4ρ
Recovered
from
disease
0.6ρ
Dead
from
disease
\newpage
This was treated sympathetically for parts (i) to (iii) as the rate of first-time sickness can be derived from this model. However in part (iv) the probability of remaining alive for three years depends on the rate of death from causes other than the disease, even among islanders who have never suffered from the disease, so this needs to be introduced. Moreover, the rate
of death from causes other than the disease may vary for persons who have never had the
disease and persons who have recovered from the disease.
Few candidates attempted part (iv). Candidates were expected to calculate the probability of
being alive in three years’ time for a person who had never had the disease at the start of that three-year period, not the probability that a person who had not had the disease would be alive and still not have had the disease in three years’ time. In part (iv) some candidates
assumed constant transition intensities and evaluate the integral, which was given full credit.
In part (v) many candidates stated that what is required is the number of healthy persons at the beginning of the period (three years ago) and the current number of healthy persons, together with the number of persons who had fallen sick. This would allow an approximate sickness rate to be calculate using various assumptions and was given partial
credit.
\end{itemize}
\end{document}

%%Page 20 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% – April 2015 – Examiners’ Report
