\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
An airline is analysing the punctuality of its scheduled flights. It measures departures
and arrivals and classifies them as early, on time or late. From its records no flights
depart early and 85% depart on time. Arrivals are early 10% of the time and late 20%
of the time.
(i)
Determine the probability that a flight arrives on time.

Further examination shows that:
•
•
5
none of the flights that depart late are early arrivals.
10% of the flights that depart on time are late arrivals.
(ii) Determine the probability that a flight will arrive early if its departure is on
time.
(iii) Show that the probability that a flight both departs on time and arrives on time
is 0.665.

(iv) Determine the probability that a flight will arrive on time if it departs late. 
(v)
 Determine the probability that if a flight arrives late it departed on time. 
[Total 11]
Q4
Let D O = departs on time, D L = departs late.
Let A E = arrives early, A O = arrives on time and A L = arrives late.
(i) P ( A O ) \;=\; 1  0.1  0.2 \;=\; 0.7 
(ii) Told P ( A E | D L ) \;=\; 0 so P ( A E  D O ) \;=\; P ( A E ) \;=\; 0.1 
P ( A E | D O ) \;=\;
P ( A E  D O )
P ( D 0 )
[Or P ( D O | A E ) \;=\;
\;=\; 0.1/ 0.85 \;=\; 0.1176
P ( A E  D O )
P ( A E )
\;=\; 1
P ( A E | D O ) \;=\; P ( D O | A E ) P ( A E ) / P ( D O ) \;=\; 1*0.1/ 0.85 \;=\; 0.1176 ]
Page 4
Subject CT3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(iii)
(
)
(
) ( )
P A L Ç D O = P A L |D O P D O = 0.1*0.85 = 0.085

P ( A O  D O ) \;=\; P ( D O )  P ( A L  D O )  P ( A E  D O ) \;=\; 0.85  0.085  0.1
\;=\; 0.665

(iv)
(v)
(
)
( ) (
)
P A O Ç D L = P A O - P A O Ç D O = 0.7 - 0.665 = 0.035 
P ( A O | D L ) \;=\; P ( A O  D L ) / P ( D L ) \;=\; 0.035 / 0.15 \;=\; 0.233 
P ( D O | A L ) \;=\; P ( A L  D O ) / P ( A L ) \;=\; 0.085 / 0.2 \;=\; 0.425 
Part (i) was answered correctly by the vast majority of candidates. The
rest of the question was not well answered. Most candidates showed
understanding of the probabilities in the context required here, but
many struggled in applying the conditional forms correctly throughout.

