\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}


\item An insurance company calculates car insurance premiums based on the age of the policyholder according to three age groups: Group A consists of drivers younger than 22 years old; Group B consists of drivers 22–33 years old; and, Group C consists of
drivers older than 33 years.

\begin{itemize}
\item Its portfolio consists of 10\% Group A policyholders, 38\% Group B policyholders and 52\% Group C policyholders.
\item 
The probability of a claim in any 12-month period for a policyholder belonging to Group A, B or C is 13\%, 3\% and 2\%, respectively.
\end{itemize}

\begin{enumerate}[(i)]
\item %(i)
Calculate the probability that a randomly chosen policyholder from this
portfolio will make a claim during a 12-month period.

One of the company’s policyholders has just made a claim.
\item %(ii)
Calculate the probability that the policyholder is younger than 22 years.
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%5
Q4
Denote by A, B, C the event that policyholder belongs to the corresponding group.
Also let F be the event that a policyholder makes a claim.
(i)

\begin{eqnarray*}
P(F) &=& P(F|A)P(A) + P(F|B)P(B) + P(F|C)P(C)\\
&=& 0.13*0.1 + 0.03*0.38 + 0.02*0.52 \\
&=& 0.0348\\
\end{eqnarray*}

P  A | F  
(ii)
P (F | A) P  A 
P  F 

0.13*0.1
 0.374
0.0348

Generally very well answered, with no particular issues.
n
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
