\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}



[Total 4]
A regulator wishes to inspect a sample of an insurer’s claims. The insurer estimates that 10\% of policies have had one claim in the last year and no policies had more than
one claim. All policies are assumed to be independent.
\begin{enumerate}
\item (i)
Determine the number of policies that the regulator would expect to examine before finding 5 claims.

On inspecting the sample claims, the regulator finds that actual payments exceeded
initial estimates by the following amounts:
\$35
(ii)
\$120
\$48
\$200
\$76
\item Find the mean and variance of these extra amounts.

It is assumed that these amounts follow a gamma distribution with parameters α
and \lambda.
(iii)
CT3 A2013–3
\item Estimate these parameters using the method of moments.

\end{enumerate}

%%-- Page 4  – %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
7
\begin{itemize}
\item (i) No of inspected policies ~ Negative binomial(5, 0.1).
Expected no of inspected policies = 5/0.1 =50
\item (ii) ∑ x = 479, ∑ x 2 = 63705
Mean = 479/5 = 95.8
Variance = (63705−5*95.8 2 )/4 = 4454.2
\item (iii)
α
α
= 95.8, V [ X ] = 2 = 4454.2
\lambda
\lambda
E [ X ]
95.8
⇒\lambda=
=
= 0.0215
V [ X ] 4454.2
E [ X ] =
⇒ α = \lambda E [ X ] = 0.0215*95.8 = 2.06
Generally very well answered with no particular issues.
\end{itemize}
\end{document}
