[Total 8]7
(i)
List EIGHT factors which should be considered when deciding whether a
model is suitable for a particular purpose.
[4]
A colleague has been asked to present a model which might be used to determine the
number of new schools required throughout a country over the next 40 years. He
forgot all about it until the last minute when he was reading an article in a newspaper
about immigration and education which provided some figures to back up the article.
Your colleague has the following suggestion for a model:
x Start with the number of children in the education system over the last twenty
years (as provided by the country’s central statistical office). Project these
forward using a straight line approach.
x Use the number of immigrants predicted to arrive in each of the next five years as
given in the newspaper article. Apply to this an estimate of “number and age of
children for each immigrant” also provided by the newspaper. Project this
forward also using a straight line approach.
x Add the two together to get the total number of children in the education system
for the next 40 years.
(ii)
CT4 S2016–5
Assess whether this model is suitable with regards to SIX of the factors which
you listed in your answer to part (i).
[6]
[Total 10]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Q7
(i)
The objectives of the modelling exercise 
The validity of the model for the purpose to which it is to be put 
The validity of the data to be used 
The validity of the assumptions used 
The possible errors associated with the model OR the fact that the parameters
used are not a perfect representation of the real world situation being modelled 
The impact of correlations between the random variables (or input variables) that
“drive” the model

The extent of correlations between the various results produced from the model 
The current relevance of models written and used in the past 
The credibility of the data input. 
The credibility of the results output 
The dangers of spurious accuracy 
The costs of buying or constructing, and of running the model 
Ease of use and availability of suitable staff to use it 
The risk of the model being used incorrectly or with wrong inputs 
The ease with which the model and its results can be communicated 
Compliance with the relevant regulations 
The existence of clear documentation
6
[Max 4]
(ii)
The objectives of the modelling exercise
The validity of the model for the purpose to which it is to be put
The model is not hugely valid as it does not address the number of schools directly,
for example by dividing the number of pupils by average school size or considering
when existing schools may become obsolete, the presence of competition etc.
[1]
The validity of the data to be used
The Central Statistical Office data will be fine, but that gained from the newspaper
will be of limited validity as estimates of future migrants arriving may be heavily
skewed by the political bias of the newspaper. Estimates of birth rates and migration
rates are generally valid data for this exercise.
[1]
%% ---  Page  9Subject CT4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5 –September 2016 – 
The validity of the assumptions used
Straight line projection is dubious over 40 years, especially on immigration numbers.
[1]
The possible errors associated with the model OR the fact that the parameters
used are not a perfect representation of the real world situation being modelled
The total number of school children in 40 years’ time is very susceptible to errors in
the parameters, for example the difference between straight line projection following a
baby boom will give a rapidly increasing number, whereas if the baby boom is over,
the numbers may decline.
[1]
The impact of correlations between the random variables that “drive” the model.
It is quite likely that the estimate of new arrivals and the children per household of
new arrivals will be biased in the same direction, i.e. both overstated or understated.
[1]
The extent of correlations between the various results produced from the model.
If you overestimate the number of children in the education system in, say 5 years’
time, you will most likely overestimate the number of children in, say, 30 years’ time
as these latter will be the next generation, the children of those in the system
in 5 years’ time.
[1]
The current relevance of models written and used in the past
The government/local authorities should have models which are still relevant even if
they need parameters adjusting.
[1]
The credibility of the data input
The data from the newspaper may be of doubtful credibility. It would be worth
examining them in the light of past trends to see whether they fall within the range of
past data.
[1]
The credibility of the results output
This model will give a very crude answer which is pretty difficult to have much faith
in. Again, it will be worth examining the output in the light of recently past trends to
see whether they mark a break with the past.
[1]
The dangers of spurious accuracy
There is no point calculating the number of children to many significant figures when
the assumptions are so approximate and the size of individual schools so variable. [1]
The cost of buying or constructing, and of running the model
An advantage of the model is that it is very inexpensive. [1]
Ease of use and availability of suitable staff to use it
The model is very easy to use. [1]
The risk of the model being used incorrectly or with wrong inputs
This is low, as the model is so simple. [1]
%% ---  Page  10Subject CT4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5 – September 2016 – 
The ease with which the model and its results can be communicated
Another advantage of the model is that it is very simple to communicate.
[1]
Compliance with the relevant regulations
Regulations are unlikely to be applicable in this case. However changes in legislation
concerning immigration might be an issue.
[1]
The existence of clear documentation
It should be easy to produce clear documentation.
[1]
[Max 6]
[Total 10]
Part (i) of this question was well answered, with many candidates scoring full
marks. Only eight factors were required for full credit. The list above gives
the complete range of factors that were awarded marks. Answers to part (ii)
were more variable. The most able candidates produced thoughtful
comments demonstrating engagement with the particular application
described in the question, and a substantial minority of candidates scored
close to full marks. Credit was given for sensible comments other than those
listed above. On the other hand, some candidates wrote only general
comments which added little to part (i).
